---
params:
  species: NA
  domain: NA
  plot_dir: NA
  emis_yr_from: NA
  emis_yr_to: NA
  folname_from: NA
  folname_to: NA
  l_fname_from: NA
  l_fname_to: NA
  l_data_from: NA
  l_data_to: NA
  uk_plot1: NA
  uk_plot2: NA
  uk_plot7: NA
title: "EMEP4UK input files: comparison report"
author: "Sam Tomlinson"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    theme: flatly
    df_print: paged
    code_folding: hide
    self_contained: true
---

```{r setup, include=FALSE}
# This Rmd is designed for HTML output only. It avoids LaTeX/PDF commands.
# The render call in compare_functions_html_v5.R sets knit_root_dir = getwd(),
# so relative paths are resolved from the project root rather than from the
# comparisons/ folder.

project_root <- knitr::opts_knit$get("root.dir")
if (is.null(project_root) || is.na(project_root) || !nzchar(project_root)) {
  project_root <- getwd()
}
project_root <- normalizePath(project_root, mustWork = FALSE)

knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center",
  out.width = "100%"
)

suppressPackageStartupMessages({
  library(data.table)
  library(knitr)
})

`%||%` <- function(x, y) {
  if (is.null(x) || length(x) == 0 || all(is.na(x))) y else x
}

q_is_abs_path <- function(path) {
  grepl("^([A-Za-z]:[\\\\/]|/|\\\\\\\\)", path)
}

q_norm_slash <- function(path) {
  gsub("\\\\", "/", path)
}

q_abs_path <- function(path) {
  if (is.null(path) || length(path) == 0 || all(is.na(path))) {
    return(NA_character_)
  }

  path <- as.character(path[[1]])
  path <- q_norm_slash(path)

  if (q_is_abs_path(path)) {
    return(normalizePath(path, winslash = "/", mustWork = FALSE))
  }

  candidates <- c(
    file.path(project_root, path),
    path
  )

  existing <- candidates[file.exists(candidates)]
  if (length(existing) > 0) {
    return(normalizePath(existing[[1]], winslash = "/", mustWork = FALSE))
  }

  normalizePath(file.path(project_root, path), winslash = "/", mustWork = FALSE)
}

plot_dir_abs <- q_abs_path(params$plot_dir)
report_dir <- normalizePath(file.path(plot_dir_abs, ".."), winslash = "/", mustWork = FALSE)

# The HTML report is written into report_dir and the PNGs are in report_dir/plots.
# Returning "plots/<filename>.png" is therefore the most robust path for browser use.
q_plot_rel <- function(path) {
  if (is.null(path) || length(path) == 0 || all(is.na(path))) return(NA_character_)

  vapply(as.character(path), function(p) {
    p <- q_norm_slash(p)
    expected <- file.path(report_dir, "plots", basename(p))

    if (file.exists(expected)) {
      return(q_norm_slash(file.path("plots", basename(p))))
    }

    p_abs <- q_abs_path(p)
    if (file.exists(p_abs) && requireNamespace("xfun", quietly = TRUE)) {
      rel <- xfun::relative_path(
        normalizePath(p_abs, winslash = "/", mustWork = TRUE),
        normalizePath(report_dir, winslash = "/", mustWork = FALSE)
      )
      return(q_norm_slash(rel))
    }

    # Best fallback: this is where the comparison functions save the PNGs.
    q_norm_slash(file.path("plots", basename(p)))
  }, character(1))
}


q_html_escape <- function(x) {
  x <- as.character(x)
  x[is.na(x)] <- ""
  x <- gsub("&", "&amp;", x, fixed = TRUE)
  x <- gsub("<", "&lt;", x, fixed = TRUE)
  x <- gsub(">", "&gt;", x, fixed = TRUE)
  x <- gsub('"', "&quot;", x, fixed = TRUE)
  x
}

q_include_plot <- function(path) {
  paths <- as.character(unlist(path, use.names = FALSE))
  paths <- paths[!is.na(paths) & nzchar(paths)]

  if (length(paths) == 0) {
    return(knitr::asis_output("<p><em>No image path was supplied.</em></p>"))
  }

  blocks <- vapply(paths, function(p) {
    full <- q_abs_path(p)

    if (!file.exists(full)) {
      # Fallback for plots saved beside the final HTML as report_dir/plots/<basename>.
      full2 <- file.path(report_dir, "plots", basename(p))
      if (file.exists(full2)) {
        full <- normalizePath(full2, winslash = "/", mustWork = TRUE)
      }
    }

    if (!file.exists(full)) {
      return(paste0(
        "<p class=\"missing-plot\"><strong>Missing plot:</strong> ",
        q_html_escape(p),
        "</p>"
      ))
    }

    # Embed the PNG as a data URI so the final HTML can be emailed/copied alone.
    img_src <- knitr::image_uri(full)

    paste0(
      "<figure class=\"plot-figure\">",
      "<img class=\"report-plot\" src=\"", q_html_escape(img_src), "\" alt=\"", q_html_escape(basename(full)), "\">",
      "<figcaption>", q_html_escape(basename(full)), "</figcaption>",
      "</figure>"
    )
  }, character(1))

  wrapper_class <- if (length(blocks) > 1) "plot-grid" else "plot-single"
  knitr::asis_output(paste0(
    "<div class=\"", wrapper_class, "\">",
    paste0(blocks, collapse = "
"),
    "</div>"
  ))
}

q_meta_val <- function(x, col) {
  dt <- x[["metadata"]]
  if (is.null(dt) || !(col %in% names(dt))) return(NA_character_)
  as.character(dt[[col]][1])
}

q_first_existing_col <- function(dt, candidates) {
  x <- candidates[candidates %in% names(dt)]
  if (length(x) == 0) return(NULL)
  x[[1]]
}

q_emis_col <- function(dt) {
  col <- q_first_existing_col(
    dt,
    c("emis_t_tot_ncoutput", "emis_t", "emis_kt", "emis", "value", "total")
  )
  if (is.null(col)) {
    stop("Could not find an emissions column in the sector table.")
  }
  col
}

q_prepare_sector_table <- function(dt) {
  dt <- as.data.table(copy(dt))

  area_col <- q_first_existing_col(
    dt,
    c("iso_char", "Area", "area", "ISO_char", "ISO", "iso")
  )
  sec_col <- q_first_existing_col(
    dt,
    c("sec_name", "GNFR", "GNFRlong", "sector_name", "sector", "sec", "SNAP")
  )
  emis_col <- q_emis_col(dt)

  if (is.null(area_col)) {
    dt[, area__ := "All"]
  } else {
    dt[, area__ := as.character(get(area_col))]
  }

  if (is.null(sec_col)) {
    dt[, sector__ := "All sectors"]
  } else {
    dt[, sector__ := as.character(get(sec_col))]
  }

  # If the source table is already in kt, keep it. Otherwise convert tonnes to kt.
  if (emis_col == "emis_kt") {
    dt_out <- dt[, .(emis_kt = sum(get(emis_col), na.rm = TRUE)), by = .(area__, sector__)]
  } else {
    dt_out <- dt[, .(emis_kt = sum(get(emis_col), na.rm = TRUE) / 1000), by = .(area__, sector__)]
  }

  setnames(dt_out, c("area__", "sector__"), c("Area", "Sector"))
  dt_out[]
}

q_safe_ratio <- function(num, den) {
  out <- rep(NA_real_, length(num))
  ok <- !is.na(num) & !is.na(den) & den != 0
  out[ok] <- num[ok] / den[ok]
  out
}

q_make_total_table <- function() {
  from <- q_prepare_sector_table(params$l_data_from[["sector_table"]])[
    , .(File_from_kt = sum(emis_kt, na.rm = TRUE)), by = Area
  ]
  to <- q_prepare_sector_table(params$l_data_to[["sector_table"]])[
    , .(File_to_kt = sum(emis_kt, na.rm = TRUE)), by = Area
  ]

  dt <- merge(from, to, by = "Area", all = TRUE)
  dt[is.na(File_from_kt), File_from_kt := 0]
  dt[is.na(File_to_kt), File_to_kt := 0]
  dt[, diff_kt := File_to_kt - File_from_kt]
  dt[, ratio_to_from := q_safe_ratio(File_to_kt, File_from_kt)]
  dt[, pct_change := (ratio_to_from - 1) * 100]

  total <- dt[, .(
    Area = "Total",
    File_from_kt = sum(File_from_kt, na.rm = TRUE),
    File_to_kt = sum(File_to_kt, na.rm = TRUE)
  )]
  total[, diff_kt := File_to_kt - File_from_kt]
  total[, ratio_to_from := q_safe_ratio(File_to_kt, File_from_kt)]
  total[, pct_change := (ratio_to_from - 1) * 100]

  dt <- rbindlist(list(dt[Area != "Total"], total), use.names = TRUE, fill = TRUE)
  dt[, sort_order__ := fifelse(Area == "Total", 1L, 0L)]
  setorder(dt, sort_order__, Area)
  dt[, sort_order__ := NULL]

  num_cols <- names(dt)[vapply(dt, is.numeric, logical(1))]
  dt[, (num_cols) := lapply(.SD, round, 3), .SDcols = num_cols]
  dt[]
}

q_make_sector_table <- function() {
  from <- q_prepare_sector_table(params$l_data_from[["sector_table"]])
  to <- q_prepare_sector_table(params$l_data_to[["sector_table"]])

  setnames(from, "emis_kt", "File_from_kt")
  setnames(to, "emis_kt", "File_to_kt")

  dt <- merge(from, to, by = c("Area", "Sector"), all = TRUE)
  dt[is.na(File_from_kt), File_from_kt := 0]
  dt[is.na(File_to_kt), File_to_kt := 0]
  dt[, diff_kt := File_to_kt - File_from_kt]
  dt[, ratio_to_from := q_safe_ratio(File_to_kt, File_from_kt)]
  dt[, pct_change := (ratio_to_from - 1) * 100]
  dt[, abs_diff_kt := abs(diff_kt)]
  setorder(dt, Area, Sector)

  num_cols <- names(dt)[vapply(dt, is.numeric, logical(1))]
  dt[, (num_cols) := lapply(.SD, round, 3), .SDcols = num_cols]
  dt[]
}

q_table <- function(dt, caption = NULL, height = "430px", page_length = 25) {
  dt <- as.data.frame(dt, stringsAsFactors = FALSE)
  if (nrow(dt) == 0) {
    return(knitr::asis_output("<p><em>No rows to display.</em></p>"))
  }

  col_headers <- paste0(
    "<th>",
    q_html_escape(names(dt)),
    "</th>",
    collapse = ""
  )

  body_rows <- apply(dt, 1, function(row) {
    paste0(
      "<tr>",
      paste0("<td>", q_html_escape(row), "</td>", collapse = ""),
      "</tr>"
    )
  })

  caption_html <- ""
  if (!is.null(caption) && length(caption) > 0 && !is.na(caption)) {
    caption_html <- paste0("<caption>", q_html_escape(caption), "</caption>")
  }

  html <- paste0(
    "<div class=\"table-scroll\" style=\"max-height:", height, ";\">",
    "<table class=\"table table-striped table-hover table-condensed report-table\">",
    caption_html,
    "<thead><tr>", col_headers, "</tr></thead>",
    "<tbody>", paste0(body_rows, collapse = "\n"), "</tbody>",
    "</table>",
    "</div>"
  )

  knitr::asis_output(html)
}

q_table_sortable <- function(dt, caption = NULL, height = "520px") {
  dt <- as.data.frame(dt, stringsAsFactors = FALSE)

  if (nrow(dt) == 0) {
    return(knitr::asis_output("<p><em>No rows to display.</em></p>"))
  }

  x <- DT::datatable(
    dt,
    caption = caption,
    rownames = FALSE,
    filter = "none",
    class = "stripe hover compact cell-border",
    options = list(
      ordering = TRUE,
      paging = FALSE,
      searching = FALSE,
      info = FALSE,
      dom = "t",
      autoWidth = TRUE
    )
  )

  htmltools::tags$div(
    class = "dt-sort-scroll",
    style = paste0("max-height:", height, ";"),
    x
  )
}

```

```{css, echo=FALSE}
body {
  font-size: 15px;
  line-height: 1.5;
}
.main-container {
  max-width: 1500px !important;
}
h1, h2, h3 {
  font-weight: 750;
}
blockquote {
  background: #f6f8fb;
  border-left: 5px solid #2c7fb8;
  padding: 0.9rem 1.1rem;
  border-radius: 0.4rem;
  margin: 1rem 0 1.4rem 0;
}
.table caption {
  caption-side: top;
  text-align: left;
  font-weight: 700;
  color: #1f2933;
}
.table:not(.dataTable) {
  width: auto !important;
  white-space: nowrap;
}
img {
  max-width: 100%;
  height: auto;
  border: 1px solid #d9e2ec;
  border-radius: 0.65rem;
  box-shadow: 0 1px 5px rgba(0,0,0,0.08);
  background: white;
  margin-bottom: 1.1rem;
}

.table-scroll {
  overflow: auto;
  width: 100%;
  border: 1px solid #d9e2ec;
  border-radius: 0.45rem;
  margin: 0.75rem 0 1.4rem 0;
}
.report-table {
  margin-bottom: 0 !important;
  white-space: nowrap;
}
.report-table caption {
  caption-side: top;
  text-align: left;
  font-weight: 700;
  color: #1f2933;
  padding: 0.6rem 0.75rem;
}
.report-table th {
  position: sticky;
  top: 0;
  background: #f8fafc;
  z-index: 2;
}
.plot-single, .plot-grid {
  margin: 1rem 0 1.6rem 0;
}
.plot-grid {
  display: grid;
  grid-template-columns: repeat(auto-fit, minmax(420px, 1fr));
  gap: 1rem;
  align-items: start;
}
.plot-figure {
  margin: 0;
  padding: 0.75rem;
  border: 1px solid #d9e2ec;
  border-radius: 0.5rem;
  background: #fff;
}
.report-plot {
  display: block;
  width: 100%;
  height: auto;
}
.plot-figure figcaption {
  margin-top: 0.45rem;
  font-size: 0.85rem;
  color: #52606d;
  word-break: break-all;
}
.missing-plot {
  color: #9b1c1c;
  background: #fff5f5;
  border: 1px solid #f8b4b4;
  border-radius: 0.4rem;
  padding: 0.75rem;
}
/* Sortable DT tables: use one outer scroll box */
.dt-sort-scroll {
  overflow: auto;
  width: 100%;
  border: 1px solid #d9e2ec;
  border-radius: 0.45rem;
  margin: 0.75rem 0 1.4rem 0;
}

.dt-sort-scroll table.dataTable {
  width: max-content !important;
  min-width: 100%;
  margin: 0 !important;
  white-space: nowrap;
}

.dt-sort-scroll table.dataTable thead th {
  position: sticky;
  top: 0;
  background: #f8fafc;
  z-index: 3;
}

.dt-sort-scroll .dataTables_wrapper {
  width: 100%;
}

.dt-sort-scroll .dataTables_info,
.dt-sort-scroll .dataTables_filter,
.dt-sort-scroll .dataTables_length {
  display: none;
}
```

# `r params$domain` input file comparison

> Comparisons are made from **File_from** to **File_to**. Positive absolute differences therefore mean **File_to is larger than File_from**. Ratios are calculated as `File_to / File_from`.

## Metadata

```{r metadata-summary}
dt_report <- data.table(
  Field = c(
    "Pollutant",
    "Domain",
    "Comparison direction",
    "Output directory",
    "Plot directory",
    "Rendered from project root"
  ),
  Value = c(
    params$species,
    params$domain,
    "File_from -> File_to",
    q_norm_slash(report_dir),
    q_norm_slash(plot_dir_abs),
    q_norm_slash(project_root)
  )
)
q_table(dt_report, caption = "Report summary", height = "260px", page_length = 20)
```

```{r metadata-inputs}
dt_meta <- data.table(
  Variable = c(
    "Project",
    "Scenario",
    "Pollutant",
    "Data source",
    "Emission year",
    "Mapping year",
    "Inventory year",
    "EMEP version",
    "NetCDF time steps",
    "Comparison time resolution",
    "Resolution",
    "Pre-2025 file format",
    "Pre-computed raster files found"
  ),
  File_from = c(
    q_meta_val(params$l_fname_from, "projectName"),
    q_meta_val(params$l_fname_from, "projectScen"),
    params$species,
    q_meta_val(params$l_fname_from, "data_source"),
    q_meta_val(params$l_fname_from, "emis_y"),
    q_meta_val(params$l_fname_from, "map_y"),
    q_meta_val(params$l_fname_from, "inv_y"),
    q_meta_val(params$l_fname_from, "EMEP_version"),
    q_meta_val(params$l_fname_from, "time_nc"),
    q_meta_val(params$l_fname_from, "time_comp"),
    q_meta_val(params$l_fname_from, "res"),
    q_meta_val(params$l_fname_from, "pre25"),
    q_meta_val(params$l_fname_from, "rast_exists")
  ),
  File_to = c(
    q_meta_val(params$l_fname_to, "projectName"),
    q_meta_val(params$l_fname_to, "projectScen"),
    params$species,
    q_meta_val(params$l_fname_to, "data_source"),
    q_meta_val(params$l_fname_to, "emis_y"),
    q_meta_val(params$l_fname_to, "map_y"),
    q_meta_val(params$l_fname_to, "inv_y"),
    q_meta_val(params$l_fname_to, "EMEP_version"),
    q_meta_val(params$l_fname_to, "time_nc"),
    q_meta_val(params$l_fname_to, "time_comp"),
    q_meta_val(params$l_fname_to, "res"),
    q_meta_val(params$l_fname_to, "pre25"),
    q_meta_val(params$l_fname_to, "rast_exists")
  )
)
q_table(dt_meta, caption = "Metadata of compared inputs", height = "360px", page_length = 30)
```

```{r metadata-paths}
dt_paths <- data.table(
  Item = c("Input folder", "NetCDF filename"),
  File_from = c(params$folname_from, paste(params$l_fname_from[["filename"]], collapse = "; ")),
  File_to = c(params$folname_to, paste(params$l_fname_to[["filename"]], collapse = "; "))
)
q_table(dt_paths, caption = "Input folders and files", height = "230px", page_length = 10)
```

## Total annual emissions map

```{r total-map, results='asis'}
q_include_plot(params$uk_plot1)
```

## Annual emissions by area

```{r total-table}
dt_total <- q_make_total_table()
q_table(
  dt_total,
  caption = paste0("Annual emissions of ", params$species, " by area. Units are kt yr-1."),
  height = "420px"
)
```

## Dominant sector contributing to change

> The positive panel shows the sector with the largest positive cell-level change. The negative panel shows the sector with the largest negative cell-level change.

```{r dominant-sector-map, results='asis'}
q_include_plot(params$uk_plot7)
```

## Largest sectoral changes

```{r largest-sector-table}
dt_sector <- q_make_sector_table()
dt_largest <- copy(dt_sector)
setorder(dt_largest, -abs_diff_kt)
dt_largest <- dt_largest[seq_len(min(nrow(dt_largest), 30))]
dt_largest[, abs_diff_kt := NULL]
q_table(
  dt_largest,
  caption = paste0("Largest absolute sectoral changes for ", params$species, ". Units are kt yr-1."),
  height = "520px"
)
```

## Sectoral maps

> The sectoral PNGs are shown directly from the `plots/` folder beside this HTML file. Open the image in a new browser tab if you need the full-size version.

```{r sector-map-grid, results='asis'}
sector_paths <- unlist(params$uk_plot2, use.names = FALSE)
sector_paths <- sector_paths[!is.na(sector_paths) & nzchar(sector_paths)]
q_include_plot(sector_paths)
```

## Annex: sectoral emissions table

```{r sector-table}
dt_sector_annex <- copy(dt_sector)
dt_sector_annex[, abs_diff_kt := NULL]
#q_table(
#  dt_sector_annex,
#  caption = paste0("Annual emissions of ", params$species, " by area and sector. Units are kt yr-1."),
#  height = "620px",
#  page_length = 50
#)

q_table_sortable(
  dt_sector_annex,
  caption = paste0("Annual emissions of ", params$species, " by area and sector. Units are kt yr-1."),
  height = "520px"
)
```

## Notes

- This HTML report avoids LaTeX-only commands and does not require a TeX installation.
- Tables are static HTML tables inside scroll boxes, rather than `DT` tables with filter boxes.
- The report is rendered with `self_contained: true`, and the PNG plots are embedded as base64 data URIs. The generated `.html` file can therefore be copied or emailed on its own, although it may be large.
