---
params:
 species: NA
 domain: NA
 plot_dir: NA
 emis_yr_from: NA
 emis_yr_to: NA
 folname_from: NA
 folname_to: NA
 l_fname_from: NA
 l_fname_to: NA
 l_data_from: NA
 l_data_to: NA
 uk_plot1: NA
 uk_plot2: NA 
 uk_plot7: NA
title: "EMEP4UK input files: comparison document"
author: "Sam Tomlinson"
date: "`r Sys.Date()`"


header-includes: 
 \usepackage{geometry}
 \geometry{top=1cm,left=1cm,bottom=1cm,right=1cm}
 
output:
  pdf_document:
     fig_caption: true
     extra_dependencies: ["threeparttable", "booktabs"]
geometry: margin=1cm
---

```{r setup, include = FALSE, echo = FALSE}
knitr::opts_knit$set(root.dir = r"[/gws/ssde/j25b/ceh_generic/samtom/EMEP_inputs]")
```

# `r params$domain` Input Files

Comparisons are made from 'File_to' perspective.

## METADATA

**Git Repo:** https://github.com/oSamTo/EMEP_inputs

**JASMIN dir:** /gws/ssde/j25b/ceh_generic/samtom/EMEP_inputs

**Output dir:** `r gsub("plots", "", params$plot_dir)`

\vspace{\baselineskip}

**Project_from** `r params$l_fname_from[["metadata"]][,projectName]`

**Folder_from:** `r params$folname_from`

**File_from:** `r params$l_fname_from[["filename"]]`

**Project_to** `r params$l_fname_to[["metadata"]][,projectName]`

**Folder_to:** `r params$folname_to`

**File_to:** `r params$l_fname_to[["filename"]]`

\vspace{\baselineskip}
```{r, message = FALSE, echo = FALSE}

dt <- data.table(Variable = c("Pollutant", "Data Source", "Emission Year", "Inventory Year", 
                              "EMEP version", "Time profiles", "Resolution"),
                 File_from = c(params$species, 
				            params$l_fname_from[["metadata"]][,data_source], 
				            params$l_fname_from[["metadata"]][,emis_y], 
				            params$l_fname_from[["metadata"]][,inv_y], 
							params$l_fname_from[["metadata"]][,EMEP_version], 
							params$l_fname_from[["metadata"]][,EMEP_version], 
							params$l_fname_from[["metadata"]][,res]),
				 File_to = c(params$species, 
				            params$l_fname_to[["metadata"]][,data_source], 
				            params$l_fname_to[["metadata"]][,emis_y], 
				            params$l_fname_to[["metadata"]][,inv_y], 
							params$l_fname_to[["metadata"]][,EMEP_version], 
							params$l_fname_to[["metadata"]][,EMEP_version], 
							params$l_fname_to[["metadata"]][,res]))
				 
kable(dt, align = "rcc", caption = "Metadata of inputs", 
	  booktabs = TRUE)
		 
```

Notes: 

* Emissions may not have been processed to the same sectors or with identical
methods, given the dateTime of production.

\newpage

## MAP: TOTAL EMISSION ANNUM^-1^

```{r, eval = TRUE, fig.align = "center", out.width = "0.9\\linewidth", echo = FALSE, fig.cap = paste0("UK/Eire total ",params$species," emissions annum-1.")}
#| fig-pos: "h"
knitr::include_graphics(paste0("/gws/ssde/j25b/ceh_generic/samtom/EMEP_inputs/",params$uk_plot1), rel_path = FALSE)
```

\vspace{\baselineskip}

```{r, message = FALSE, echo = FALSE, label = tab-tot}

dt_tots_from <- copy(params$l_data_from[["sector_table"]])
dt_tots_from <- dt_tots_from[, .(emis_kt = sum(emis_t_tot_ncoutput, na.rm=T)/1000), by = iso_char]

dt_tots_to <- copy(params$l_data_to[["sector_table"]])
dt_tots_to <- dt_tots_to[, .(emis_kt = sum(emis_t_tot_ncoutput, na.rm=T)/1000), by = iso_char]

dt1 <- data.table(Area = c("UK", "Eire", "SEA"),
                  Species = params$species,
 				  File_from_kt = c(dt_tots_from[iso_char == "UK", round(emis_kt, 2)],
				                dt_tots_from[iso_char == "IE", round(emis_kt, 2)],
				                dt_tots_from[iso_char == "SEA", round(emis_kt, 2)]),
				  File_to_kt = c(dt_tots_to[iso_char == "UK", round(emis_kt, 2)],
				                dt_tots_to[iso_char == "IE", round(emis_kt, 2)],
				                dt_tots_to[iso_char == "SEA", round(emis_kt, 2)]))
								
dt1[, diff_kt    := File_to_kt - File_from_kt]
dt1[, diff_alpha := round(File_to_kt / File_from_kt, 3)]
							  
dt2 <- data.table(Area = c("Total"),
                  Species = params$species,
 				  File_from_kt = dt_tots_from[, round(sum(emis_kt,na.rm=T), 2)],
				  File_to_kt = dt_tots_to[, round(sum(emis_kt,na.rm=T), 2)])

dt2[, diff_kt    := File_to_kt - File_from_kt]
dt2[, diff_alpha := round(File_to_kt / File_from_kt, 3)]

dt <- rbindlist(list(dt1, dt2), use.names = T)

kable(dt, align = "cccc", caption = paste0("Annual emissions of ", params$species, " per ISO."), 
	  booktabs = TRUE)
```
```{r, results='asis', echo=FALSE}
if (params$domain == "UKEIRE") {

cat("
* N.B. UK and Eire totals in Table 2 are only on-land emissions.
")

}
```


\newpage

## MAP: DOMINANT SECTORAL EMISSIONS CHANGE ANNUM^-1^

```{r, eval = TRUE, fig.align = "center", out.width = "0.85\\linewidth", echo = FALSE, fig.cap = paste0("UK/Eire dominant sectoral change (pos and neg) for ",params$species,".")}
#| fig-pos: "h"
knitr::include_graphics(paste0("/gws/ssde/j25b/ceh_generic/samtom/EMEP_inputs/",params$uk_plot7), rel_path = FALSE)
```

\newpage

```{r, eval = TRUE, fig.align = "center", out.width = "42%", echo = FALSE, fig.show="hold", fig.cap = ""}
knitr::include_graphics(c(paste0("/gws/ssde/j25b/ceh_generic/samtom/EMEP_inputs/",params$uk_plot2[[1]]),
                          paste0("/gws/ssde/j25b/ceh_generic/samtom/EMEP_inputs/",params$uk_plot2[[2]])), rel_path = FALSE)
```
```{r, eval = TRUE, fig.align = "center", out.width = "42%", echo = FALSE, fig.show="hold", fig.cap = ""}
knitr::include_graphics(c(paste0("/gws/ssde/j25b/ceh_generic/samtom/EMEP_inputs/",params$uk_plot2[[3]]),
                          paste0("/gws/ssde/j25b/ceh_generic/samtom/EMEP_inputs/",params$uk_plot2[[4]])), rel_path = FALSE)
```

\newpage

```{r, eval = TRUE, fig.align = "center", out.width = "42%", echo = FALSE, fig.show="hold", fig.cap = ""}

knitr::include_graphics(c(paste0("/gws/ssde/j25b/ceh_generic/samtom/EMEP_inputs/",params$uk_plot2[[5]]),
                          paste0("/gws/ssde/j25b/ceh_generic/samtom/EMEP_inputs/",params$uk_plot2[[6]])), rel_path = FALSE)
```
```{r, eval = TRUE, fig.align = "center", out.width = "42%", echo = FALSE, fig.show="hold", fig.cap = ""}

knitr::include_graphics(c(paste0("/gws/ssde/j25b/ceh_generic/samtom/EMEP_inputs/",params$uk_plot2[[7]]),
                          paste0("/gws/ssde/j25b/ceh_generic/samtom/EMEP_inputs/",params$uk_plot2[[8]])), rel_path = FALSE)
```

\newpage

```{r, eval = TRUE, fig.align = "center", out.width = "42%", echo = FALSE, fig.show="hold", fig.cap = ""}

knitr::include_graphics(c(paste0("/gws/ssde/j25b/ceh_generic/samtom/EMEP_inputs/",params$uk_plot2[[9]]),
                          paste0("/gws/ssde/j25b/ceh_generic/samtom/EMEP_inputs/",params$uk_plot2[[10]])), rel_path = FALSE)
```
```{r, eval = TRUE, fig.align = "center", out.width = "42%", echo = FALSE, fig.show="hold", fig.cap = ""}

knitr::include_graphics(c(paste0("/gws/ssde/j25b/ceh_generic/samtom/EMEP_inputs/",params$uk_plot2[[11]]),
                          paste0("/gws/ssde/j25b/ceh_generic/samtom/EMEP_inputs/",params$uk_plot2[[12]])), rel_path = FALSE)
```

\newpage

```{r, eval = TRUE, fig.align = "center", out.width = "42%", echo = FALSE, fig.show="hold", fig.cap = ""}

knitr::include_graphics(c(paste0("/gws/ssde/j25b/ceh_generic/samtom/EMEP_inputs/",params$uk_plot2[[13]])), rel_path = FALSE)

```

\newpage

## ANNEX TABLE: SECTORAL EMISSIONS ANNUM^-1^

```{r, message = FALSE, echo = FALSE, label = tab-sec-tot}

dt_tots_from <- copy(params$l_data_from[["sector_table"]])
layer_name_from <- params$l_fname_from$layer_name
dt_tots_from <- dt_tots_from[, .(emis_kt = round(sum(emis_t_tot_ncoutput, na.rm=T)/1000, 3)), by = .(iso_char, sec_name)][iso_char != "SEA"]
setnames(dt_tots_from, "emis_kt", layer_name_from)

dt_tots_to <- copy(params$l_data_to[["sector_table"]])
layer_name_to <- params$l_fname_to$layer_name
dt_tots_to <- dt_tots_to[, .(emis_kt = round(sum(emis_t_tot_ncoutput, na.rm=T)/1000, 3)), by = .(iso_char, sec_name)][iso_char != "SEA"]
setnames(dt_tots_to, "emis_kt", layer_name_to)

dt <- dt_tots_from[dt_tots_to, on = c("iso_char", "sec_name")]
								
dt[, diff_kt    := get(layer_name_to) - get(layer_name_from)]
dt[, diff_alpha := round(get(layer_name_to) / get(layer_name_from), 3)]
	

kable(dt, align = "crcccc", caption = paste0("Annual emissions of ", params$species, " per sector per ISO."), 
	  booktabs = TRUE)
```
