---
params:
 species: NA
 plot_dir: NA
 emis_yr_1: NA
 emis_yr_2: NA
 folname_1: NA
 folname_2: NA
 l_fname_1: NA
 l_fname_2: NA
 l_m_1: NA
 l_m_2: NA
 uk_plot1: NA
 uk_plot2: NA
 uk_plot3: NA
 uk_plot4: NA
 uk_plot5: NA
 uk_plot6: NA
title: "EMEP4UK input files Comparison document - UK/EIRE & EU"
author: "Sam Tomlinson"
date: "`r Sys.Date()`"


header-includes: 
 \usepackage{geometry}
 \geometry{top=1cm,left=1cm,bottom=1cm,right=1cm}
 
output:
  pdf_document:
     fig_caption: true
     extra_dependencies: ["threeparttable", "booktabs"]
geometry: margin=1cm
---

```{r setup, include = FALSE, echo = FALSE}
knitr::opts_knit$set(root.dir = r"[/gws/nopw/j04/ceh_generic/samtom/EMEP_inputs]")
```

# UK/EIRE INPUT FILES

Comparisons are made from File_2 perspective.

## METADATA

**Project:** NFC

**Git Repo:** https://github.com/oSamTo/EMEP_inputs

**JASMIN dir:** /gws/nopw/j04/ceh_generic/samtom/EMEP_inputs

**Output dir:** `r gsub("plots", "", params$plot_dir)`

\vspace{\baselineskip}

**Folder_1:** `r params$folname_1`

**File_1:** `r params$l_fname_1[["filename"]]`

**Folder_2:** `r params$folname_2`

**File_2:** `r params$l_fname_2[["filename"]]`

\vspace{\baselineskip}
```{r, message = FALSE, echo = FALSE}

dt <- data.table(Variable = c("Pollutant", "Emission Year", "Inventory Year", 
                              "EMEP version", "Time profiles", "Resolution"),
                 File_1 = c(params$species, 
				            params$l_fname_1[["metadata"]][,emis_y], 
				            params$l_fname_1[["metadata"]][,inv_y], 
							params$l_fname_1[["metadata"]][,EMEP_version], 
							params$l_fname_1[["metadata"]][,EMEP_version], 
							params$l_fname_1[["metadata"]][,res]),
				 File_2 = c(params$species, 
				            params$l_fname_2[["metadata"]][,emis_y], 
				            params$l_fname_2[["metadata"]][,inv_y], 
							params$l_fname_2[["metadata"]][,EMEP_version], 
							params$l_fname_2[["metadata"]][,EMEP_version], 
							params$l_fname_2[["metadata"]][,res]))
				 
kable(dt, align = "rcc", caption = "Metadata of inputs", 
	  booktabs = TRUE)
		 
```

Notes: 

* Emissions may not have been processed to the same sectors or with identical
methods, given the dateTime of production.

* Emissions from the UK and Eire are masked to a UK+10km mask. EMEP/CEIP 
emissions take precedence outside of this mask. 

* The 10km strip from the land territory into the sea is combined (UK + Eire) to a 'SEA' ISO 
code, for all sectors.

\newpage

## MAP: TOTAL EMISSION ANNUM^-1^

```{r, eval = TRUE, fig.align = "center", out.width = "0.9\\linewidth", echo = FALSE, fig.cap = paste0("UK/Eire total ",params$species," emissions annum-1.")}
#| fig-pos: "h"
knitr::include_graphics(paste0("/gws/nopw/j04/ceh_generic/samtom/EMEP_inputs/",params$uk_plot1), rel_path = FALSE)
```

\vspace{\baselineskip}

```{r, message = FALSE, echo = FALSE, label = tab-tot}

dt_tots_1 <- copy(params$l_m_1[["summary"]])
dt_tots_1 <- dt_tots_1[, .(emis_kt = sum(emis_t, na.rm=T)/1000), by = Area]

dt_tots_2 <- copy(params$l_m_2[["summary"]])
dt_tots_2 <- dt_tots_2[, .(emis_kt = sum(emis_t, na.rm=T)/1000), by = Area]

dt1 <- data.table(Area = c("UK", "Eire", "SEA"),
                  Species = params$species,
 				  File_1_kt = c(dt_tots_1[Area == "UK", round(emis_kt, 2)],
				                dt_tots_1[Area == "IE", round(emis_kt, 2)],
				                dt_tots_1[Area == "SEA", round(emis_kt, 2)]),
				  File_2_kt = c(dt_tots_2[Area == "UK", round(emis_kt, 2)],
				                dt_tots_2[Area == "IE", round(emis_kt, 2)],
				                dt_tots_2[Area == "SEA", round(emis_kt, 2)]))
								
dt1[, diff_kt    := File_2_kt - File_1_kt]
dt1[, diff_alpha := round(File_2_kt / File_1_kt, 3)]
							  
dt2 <- data.table(Area = c("Total"),
                  Species = params$species,
 				  File_1_kt = dt_tots_1[, round(sum(emis_kt,na.rm=T), 2)],
				  File_2_kt = dt_tots_2[, round(sum(emis_kt,na.rm=T), 2)])

dt2[, diff_kt    := File_2_kt - File_1_kt]
dt2[, diff_alpha := round(File_2_kt / File_1_kt, 3)]

dt <- rbindlist(list(dt1, dt2), use.names = T)

kable(dt, align = "cccc", caption = paste0("Annual emissions of ", params$species, " per ISO."), 
	  booktabs = TRUE)
```

* N.B. UK and Eire totals in Table 2 are only on-land emissions.

\newpage

## MAP: SECTORAL EMISSIONS ANNUM^-1^

```{r, message = FALSE, echo = FALSE, label = tab-sec-tot}

dt_tots_1 <- copy(params$l_m_1[["summary"]])
fname1 <- dt_tots_1[,unique(plot_group)]
dt_tots_1 <- dt_tots_1[, .(emis_kt = round(sum(emis_t, na.rm=T)/1000, 3)), by = .(Area,GNFR)][Area != "SEA"]
setnames(dt_tots_1, "emis_kt", fname1)

dt_tots_2 <- copy(params$l_m_2[["summary"]])
fname2 <- dt_tots_2[,unique(plot_group)]
dt_tots_2 <- dt_tots_2[, .(emis_kt = round(sum(emis_t, na.rm=T)/1000, 3)), by = .(Area,GNFR)][Area != "SEA"]
setnames(dt_tots_2, "emis_kt", fname2)

dt <- dt_tots_1[dt_tots_2, on = c("Area", "GNFR")]
								
dt[, diff_kt    := get(fname2) - get(fname1)]
dt[, diff_alpha := round(get(fname2) / get(fname1), 3)]
	

kable(dt, align = "crcccc", caption = paste0("Annual emissions of ", params$species, " per sector per ISO."), 
	  booktabs = TRUE)
```

```{r, eval = TRUE, fig.align = "center", out.width = "0.7\\linewidth", echo = FALSE, fig.cap = paste0("UK/Eire sectoral ",params$species," emissions annum-1.")}
#| fig-pos: "h"
knitr::include_graphics(paste0("/gws/nopw/j04/ceh_generic/samtom/EMEP_inputs/",params$uk_plot6), rel_path = FALSE)
```

\newpage

```{r, eval = TRUE, fig.align = "center", out.width = "42%", echo = FALSE, fig.show="hold", fig.cap = ""}
knitr::include_graphics(c(paste0("/gws/nopw/j04/ceh_generic/samtom/EMEP_inputs/",params$uk_plot2[[1]]),
                          paste0("/gws/nopw/j04/ceh_generic/samtom/EMEP_inputs/",params$uk_plot2[[2]])), rel_path = FALSE)
```
```{r, eval = TRUE, fig.align = "center", out.width = "42%", echo = FALSE, fig.show="hold", fig.cap = ""}
knitr::include_graphics(c(paste0("/gws/nopw/j04/ceh_generic/samtom/EMEP_inputs/",params$uk_plot2[[3]]),
                          paste0("/gws/nopw/j04/ceh_generic/samtom/EMEP_inputs/",params$uk_plot2[[4]])), rel_path = FALSE)
```

\newpage

```{r, eval = TRUE, fig.align = "center", out.width = "42%", echo = FALSE, fig.show="hold", fig.cap = ""}

knitr::include_graphics(c(paste0("/gws/nopw/j04/ceh_generic/samtom/EMEP_inputs/",params$uk_plot2[[5]]),
                          paste0("/gws/nopw/j04/ceh_generic/samtom/EMEP_inputs/",params$uk_plot2[[6]])), rel_path = FALSE)
```
```{r, eval = TRUE, fig.align = "center", out.width = "42%", echo = FALSE, fig.show="hold", fig.cap = ""}

knitr::include_graphics(c(paste0("/gws/nopw/j04/ceh_generic/samtom/EMEP_inputs/",params$uk_plot2[[7]]),
                          paste0("/gws/nopw/j04/ceh_generic/samtom/EMEP_inputs/",params$uk_plot2[[8]])), rel_path = FALSE)
```

\newpage

```{r, eval = TRUE, fig.align = "center", out.width = "42%", echo = FALSE, fig.show="hold", fig.cap = ""}

knitr::include_graphics(c(paste0("/gws/nopw/j04/ceh_generic/samtom/EMEP_inputs/",params$uk_plot2[[9]]),
                          paste0("/gws/nopw/j04/ceh_generic/samtom/EMEP_inputs/",params$uk_plot2[[10]])), rel_path = FALSE)
```
```{r, eval = TRUE, fig.align = "center", out.width = "42%", echo = FALSE, fig.show="hold", fig.cap = ""}

knitr::include_graphics(c(paste0("/gws/nopw/j04/ceh_generic/samtom/EMEP_inputs/",params$uk_plot2[[11]]),
                          paste0("/gws/nopw/j04/ceh_generic/samtom/EMEP_inputs/",params$uk_plot2[[12]])), rel_path = FALSE)
```

\newpage

```{r, eval = TRUE, fig.align = "center", out.width = "42%", echo = FALSE, fig.show="hold", fig.cap = ""}

knitr::include_graphics(c(paste0("/gws/nopw/j04/ceh_generic/samtom/EMEP_inputs/",params$uk_plot2[[13]])), rel_path = FALSE)

```

\newpage

## MAP: EMISSIONS MONTH^-1^

```{r, eval = TRUE, fig.align = "center", out.width = "26%", echo = FALSE, fig.show="hold", fig.cap = ""}

knitr::include_graphics(c(paste0("/gws/nopw/j04/ceh_generic/samtom/EMEP_inputs/",params$uk_plot3[[1]]),
                          paste0("/gws/nopw/j04/ceh_generic/samtom/EMEP_inputs/",params$uk_plot3[[2]]),
						  paste0("/gws/nopw/j04/ceh_generic/samtom/EMEP_inputs/",params$uk_plot3[[3]])), rel_path = FALSE)
```

```{r, eval = TRUE, fig.align = "center", out.width = "26%", echo = FALSE, fig.show="hold", fig.cap = ""}

knitr::include_graphics(c(paste0("/gws/nopw/j04/ceh_generic/samtom/EMEP_inputs/",params$uk_plot3[[4]]),
                          paste0("/gws/nopw/j04/ceh_generic/samtom/EMEP_inputs/",params$uk_plot3[[5]]),
						  paste0("/gws/nopw/j04/ceh_generic/samtom/EMEP_inputs/",params$uk_plot3[[6]])), rel_path = FALSE)
```

```{r, eval = TRUE, fig.align = "center", out.width = "26%", echo = FALSE, fig.show="hold", fig.cap = ""}

knitr::include_graphics(c(paste0("/gws/nopw/j04/ceh_generic/samtom/EMEP_inputs/",params$uk_plot3[[7]]),
                          paste0("/gws/nopw/j04/ceh_generic/samtom/EMEP_inputs/",params$uk_plot3[[8]]),
						  paste0("/gws/nopw/j04/ceh_generic/samtom/EMEP_inputs/",params$uk_plot3[[9]])), rel_path = FALSE)
```

\newpage

```{r, eval = TRUE, fig.align = "center", out.width = "28%", echo = FALSE, fig.show = "hold", fig.cap = ""}

knitr::include_graphics(c(paste0("/gws/nopw/j04/ceh_generic/samtom/EMEP_inputs/",params$uk_plot3[[10]]),
                          paste0("/gws/nopw/j04/ceh_generic/samtom/EMEP_inputs/",params$uk_plot3[[11]]),
						  paste0("/gws/nopw/j04/ceh_generic/samtom/EMEP_inputs/",params$uk_plot3[[12]])), rel_path = FALSE)
```

```{r, eval = TRUE, fig.align = "center", out.width = "0.8\\linewidth", echo = FALSE, fig.cap = paste0("UK/Eire total ",params$species," emissions month-1.")}
#| fig-pos: "h"
knitr::include_graphics(paste0("/gws/nopw/j04/ceh_generic/samtom/EMEP_inputs/",params$uk_plot4), rel_path = FALSE)
```

\newpage

```{r, eval = TRUE, fig.align = "center", out.width = "0.9\\linewidth", echo = FALSE, fig.cap = paste0("UK/Eire sectoral ",params$species," emissions month-1.")}
#| fig-pos: "h"
knitr::include_graphics(paste0("/gws/nopw/j04/ceh_generic/samtom/EMEP_inputs/",params$uk_plot5), rel_path = FALSE)
```
