---
params:
 y: NA
 dt_poll: NA
 species: NA
 uk_folname: NA
 eu_folname: NA
 map_yr_uk: NA
 naei_inv: NA
 emep_inv: NA
 v_EMEP_sec: NA
 dt_sec: NA
 time_dim: NA
 emep_version: NA
 uk_agg_schema: NA
 eu_agg_schema: NA
 tp_scheme: NA
 l_fname_uk_sums: NA
 l_fname_eu_sums: NA
 l_uk_maps : NA
 l_eu_maps : NA
 dt_month_uk: NA
 dt_month_eu: NA
title: "EMEP4UK input files QAQC document - UK/EIRE & EU"
author: "Sam Tomlinson"
date: "`r Sys.Date()`"


header-includes: 
 \usepackage{geometry}
 \geometry{top=1cm,left=1cm,bottom=1cm,right=1cm}
 
output:
  pdf_document:
     fig_caption: true
     extra_dependencies: ["threeparttable", "booktabs"]
---

# UK/EIRE INPUT FILES

## METADATA

**Project:** NFC

**Git Repo:** https://github.com/oSamTo/EMEP_inputs

**JASMIN dir:** /gws/nopw/j04/ceh_generic/samtom/EMEP_inputs

**Output dir:** `r gsub("/gws/nopw/j04/ceh_generic/samtom/EMEP_inputs/", "", params$uk_folname)`

**EMEP version:** `r params$emep_version`

\vspace{\baselineskip}

Input files processed to 0.01$^\circ$ x 0.01$^\circ$ resolution, in WGS84. 

\vspace{\baselineskip}

```{r, message = FALSE, echo = FALSE}

dt <- data.table(Area = c("UK", "Eire", "SEA"),
                 Species = params$species,
                 Inv_Year = c(params$naei_inv, params$emep_inv, "as above"),
				 Emis_Year = c(params$y, params$y, "as above"),
				 temp_res = c("annual", "annual", "N/A"),
				 orig_sec = c("SNAP", "GNFR", "N/A"),
				 orig_res = c("1km", "1km", "N/A"),
				 orig_crs = c("BNG", "TM65", "N/A"))

kable(dt, align = "cccccccc", caption = "Metadata of inputs", 
	  booktabs = TRUE)
		 
```

Notes: 

* Emissions from the UK and Eire are processed to consistent sector categories (see 
/data/lookups/EMEP_sectors.csv) and masked to a UK+10km mask. EMEP/CEIP 
emissions take precedence outside of this mask. 

* The 10km strip from the land territory into the sea is combined (UK + Eire) to a 'SEA' ISO 
code, for all sectors.

* Emissions maps are pre-processed from 1km x 1km BNG (UK) and 1km x 1km Irish 
Grid TM65 (Eire) into WGS84 lat lon, at 0.01$^\circ$ x 0.01$^\circ$ resolution. 

  * This is done in /gws/nopw/j04/ceh_generic/inventory_processor

\newpage

## MAP: TOTAL EMISSION ANNUM^-1^

```{r, eval=TRUE, fig.align="center", out.width = "0.7\\linewidth", echo = FALSE, fig.cap = paste0("UK/Eire total ",params$species," emissions annum-1 (Non-UK/Eire emissions for context).")}
#| fig-pos: "h"
knitr::include_graphics(paste0("/gws/nopw/j04/ceh_generic/samtom/EMEP_inputs/outputs/EMEP4UK",params$emep_version,"/inv",params$naei_inv,"/UKEIRE/",params$tp_scheme,"/TP",tp_scheme,"_",uk_agg_schema,"/plots/e",params$y,"/",params$species,"_UKEIRE_",params$y,"emis_",params$map_yr_uk,"map_",params$naei_inv,"inv_1_UKTOTMAP.png"))
```

\vspace{\baselineskip}

```{r, message = FALSE, echo = FALSE, label = tab-tot}

l_ann_tot <- params$l_uk_maps$ann_tot_all
s_ann_tot <- rast(l_ann_tot)
s_ann_tot[s_ann_tot == 0] <- NA

dt_tot_emis <- fread(params$l_fname_uk_sums[["ncoutput"]])

dt1 <- data.table(Area = c("UK", "Eire", "SEA"),
                  Species = params$species,
 				  Total_kt = c(dt_tot_emis[Area == "GB", round(sum(emis_t_tot_ncoutput, na.rm = T)/1000, 2)],
				               dt_tot_emis[Area == "IE", round(sum(emis_t_tot_ncoutput, na.rm = T)/1000, 2)],
				               dt_tot_emis[Area == "SEA", round(sum(emis_t_tot_ncoutput, na.rm = T)/1000, 2)]),
				  mean_t   = c(round(global(s_ann_tot[[grep("_GB", names(s_ann_tot))]], mean, na.rm=T)[,1], 2),
				               round(global(s_ann_tot[[grep("_IE", names(s_ann_tot))]], mean, na.rm=T)[,1], 2),
					 		   round(global(s_ann_tot[[grep("_SEA", names(s_ann_tot))]], mean, na.rm=T)[,1], 2)),
				  max_t    = c(round(global(s_ann_tot[[grep("_GB", names(s_ann_tot))]], max, na.rm=T)[,1], 2),
				               round(global(s_ann_tot[[grep("_IE", names(s_ann_tot))]], max, na.rm=T)[,1], 2),
					 		   round(global(s_ann_tot[[grep("_SEA", names(s_ann_tot))]], max, na.rm=T)[,1], 2)))
							  
dt2 <- data.table(Area = c("Overall"),
                  Species = params$species,
 				  Total_kt = dt1[, sum(Total_kt, na.rm=T)],
				  mean_t   = round(global(app(s_ann_tot, sum, na.rm=T), mean, na.rm=T)[,1], 2),
				  max_t    = round(global(app(s_ann_tot, sum, na.rm=T), max, na.rm=T)[,1], 2))

dt <- rbindlist(list(dt1, dt2), use.names = T)

kable(dt, align = "ccc", caption = paste0("Annual emissions of ", params$species, " per ISO."), 
	  booktabs = TRUE)
```

* N.B. UK and Eire totals in Table 2 are only on-land emissions.

\newpage

## PROCESSING FLOW

The processing in Figure \ref{fig:fig-proc1} shows total emissions as taken 
from the maps (inv_spatial), the inventory table (inv_table), the scaled maps 
(spatial_scaled), and the area aggregated totals using the mask. 

N.B. the map used is not necessarily the same as the emissions year (e.g. MapEire from 2019).

```{r, eval=TRUE, fig.align="center", out.width = "0.6\\linewidth", echo = FALSE, fig.cap = "Emissions processing from source to area-scaled totals.", label = fig-proc1}
#| fig-pos: "h"
knitr::include_graphics(paste0("/gws/nopw/j04/ceh_generic/samtom/EMEP_inputs/outputs/EMEP4UK",params$emep_version,"/inv",params$naei_inv,"/UKEIRE/",params$tp_scheme,"/TP",tp_scheme,"_",uk_agg_schema,"/plots/e",params$y,"/",params$species,"_UKEIRE_",params$y,"emis_",params$map_yr_uk,"map_",params$naei_inv,"inv_2_UKINVBAR.png"))
```

Total `r params$species` in UK & Eire, processed: `r round(fread(params$l_fname_uk_sums[["masked"]])[Area == "uk",sum(emis_t_tot_masked, na.rm=T)/1000],1)` kt + `r round(fread(params$l_fname_uk_sums[["masked"]])[Area == "ie",sum(emis_t_tot_masked, na.rm=T)/1000],1)` kt = `r round(fread(params$l_fname_uk_sums[["processed"]])[,sum(emis_t_tot_grouped, na.rm=T)/1000],1)` kt. Spatial breakdown :

* UK = `r round(fread(params$l_fname_uk_sums[["processed"]])[Area == "uk", sum(emis_t_tot_grouped, na.rm=T)/1000], 1)` kt

* Eire = `r round(fread(params$l_fname_uk_sums[["processed"]])[Area == "ie", sum(emis_t_tot_grouped, na.rm=T)/1000], 1)` kt

* SEA = `r round(fread(params$l_fname_uk_sums[["processed"]])[Area == "sea", sum(emis_t_tot_grouped, na.rm=T)/1000], 1)` kt

* Outwith 10km = `r round(fread(params$l_fname_uk_sums[["processed"]])[Area == "ow", sum(emis_t_tot_grouped, na.rm=T)/1000], 1)` kt

Figure \ref{fig:fig-proc2} shows the data progressing to NetCDF inputs - 'outwith' data (data outside 10km buffer) is dropped. 

```{r, eval=TRUE, fig.align="center", out.width = "0.6\\linewidth", echo = FALSE, fig.cap = "Area-scaled toals to NetCDF input data.", label = fig-proc2}
#| fig-pos: "h"
knitr::include_graphics(paste0("/gws/nopw/j04/ceh_generic/samtom/EMEP_inputs/outputs/EMEP4UK",params$emep_version,"/inv",params$naei_inv,"/UKEIRE/",params$tp_scheme,"/TP",tp_scheme,"_",uk_agg_schema,"/plots/e",params$y,"/",params$species,"_UKEIRE_",params$y,"emis_",params$map_yr_uk,"map_",params$naei_inv,"inv_3_UKNCBAR.png"))
```

Emissions coming from processed inventories to NetCDF inputs; 'outwith' emissions (emissions outside 10km sea buffer) are masked and not used for UK/Eire (replaced by EMEP-EU).

\newpage

## EMISSIONS SECTORAL CLASSIFICATIONS

### UK data starts in SNAP sectors, Eire data in GNFR sectors. 

* All UK SNAP03 & SNAP04 goes to B_Industry.

* All UK SNAP08 goes to I_Offroad. 

* All UK SNAP10 goes to K_AgriLivestock. 

```{r, message = FALSE, echo = FALSE}

dt <- params$dt_sec[,-1]

kable(dt, align = "rcccccl", caption = "Sectoral lookup table, SNAP <--> GNFR", 
	  booktabs = TRUE)
```

\newpage

## MAP: SECTORAL EMISSIONS ANNUM^-1^

### UK and Eire temporal schema: `r params$time_dim`

Source: `r if(params$time_dim == "annual"){paste0("EMEP ", params$emep_version, " schema: data/temporal/EMEP4UK",params$emep_version,"/")}else{paste0("UKEM ", params$tp_scheme, " schema: /gws/nopw/j04/ukem/test/sam/ukem_pro/output/")}`

```{r, eval=TRUE, fig.align="center", out.width = "0.75\\linewidth", echo = FALSE, fig.cap = paste0("UK/Eire total ",params$species," emissions sector-1 annum-1")}
#| fig-pos: "h"
knitr::include_graphics(paste0("/gws/nopw/j04/ceh_generic/samtom/EMEP_inputs/outputs/EMEP4UK",params$emep_version,"/inv",params$naei_inv,"/UKEIRE/",params$tp_scheme,"/TP",tp_scheme,"_",uk_agg_schema,"/plots/e",params$y,"/",params$species,"_UKEIRE_",params$y,"emis_",params$map_yr_uk,"map_",params$naei_inv,"inv_6_UKANNSECMAP.png"))
```

```{r, message = FALSE, echo = FALSE}

dt_out <- fread(params$l_fname_uk_sums[["ncoutput"]])

dt_long <- dt_out[,list(emis_kt = round(sum(emis_t_tot_ncoutput, na.rm=T)/1000,3)), by = .(Area, sec_name)]
dt_wide <- dcast(dt_long, sec_name~Area, value.var = "emis_kt")
dt_wide[, Total := rowSums(.SD, na.rm=T), .SDcols = c("GB", "IE", "SEA")]
dt_wide <- adorn_totals(dt_wide)
setnames(dt_wide, "sec_name", "Sector")

kable(dt_wide, align = "rccc", caption = paste0(params$species," emissions (kt) per sector per ISO"), 
	  booktabs = TRUE)
```

\newpage

## EMISSIONS MONTH^-1^

### UK and Eire temporal schema: `r params$time_dim`

Source: `r if(params$time_dim == "annual"){paste0("EMEP ", params$emep_version, " schema: data/temporal/EMEP4UK",params$emep_version,"/")}else{paste0("UKEM ", params$tp_scheme, " schema: /gws/nopw/j04/ukem/test/sam/ukem_pro/output/")}`

```{r, eval=TRUE, fig.align="center", out.width = "0.6\\linewidth", echo = FALSE, fig.cap = paste0("Emissions of ",params$species," month-1"), label = "fig-montot"}
#| fig-pos: "h"
knitr::include_graphics(paste0("/gws/nopw/j04/ceh_generic/samtom/EMEP_inputs/outputs/EMEP4UK",params$emep_version,"/inv",params$naei_inv,"/UKEIRE/",params$tp_scheme,"/TP",tp_scheme,"_",uk_agg_schema,"/plots/e",params$y,"/",params$species,"_UKEIRE_",params$y,"emis_",params$map_yr_uk,"map_",params$naei_inv,"inv_4_UKMONTOTLINE.png"))
```

Total monthly emissions of `r params$species` in UK, Eire and the SEA buffer. 
`r if(params$time_dim == "annual"){paste0("As annual emissions input into NetCDF, using model profiles.")}`
`r if(params$time_dim == "month"){paste0("Temporal profiles taken from UKEM")}`

```{r, message = FALSE, echo = FALSE}

dt <- params$dt_month_uk[,list(emis_kt = round(sum(emis_t, na.rm=T)/1000, 2)), by = .(Area, time)]
dt[, Area := toupper(Area)]
dt[, time := paste0("M", time)]
dt[, time := factor(time, levels = paste0("M",1:12))]
dt_wide <- dcast(dt, Area~time, value.var = "emis_kt")
dt_wide[, Total := rowSums(.SD, na.rm=T), .SDcols = paste0("M",1:12)]
dt_wide <- adorn_totals(dt_wide)

kable(dt_wide, align = "rccccccccccccc", caption = paste0("Total monthly ", params$species, " emissions (kt)"), 
	  booktabs = TRUE)
```

\newpage

## MAP: EMISSIONS MONTH^-1^

Also see Table 5 above for totals. 

```{r, eval=TRUE, fig.align="center", out.width = "0.85\\linewidth", echo = FALSE, fig.cap = paste0("UK/Eire total ",params$species," emissions month-1")}
#| fig-pos: "h"
knitr::include_graphics(paste0("/gws/nopw/j04/ceh_generic/samtom/EMEP_inputs/outputs/EMEP4UK",params$emep_version,"/inv",params$naei_inv,"/UKEIRE/",params$tp_scheme,"/TP",tp_scheme,"_",uk_agg_schema,"/plots/e",params$y,"/",params$species,"_UKEIRE_",params$y,"emis_",params$map_yr_uk,"map_",params$naei_inv,"inv_7_UKMONTOTMAP.png"))
```

\newpage

## SECTORAL EMISSIONS MONTH^-1^

### UK and Eire temporal schema: `r params$time_dim`

Source: `r if(params$time_dim == "annual"){paste0("EMEP ", params$emep_version, " schema: data/temporal/EMEP4UK",params$emep_version,"/")}else{paste0("UKEM ", params$tp_scheme, " schema: /gws/nopw/j04/ukem/test/sam/ukem_pro/output/")}`

```{r, eval=TRUE, fig.align="center", out.width = "0.7\\linewidth", echo = FALSE, fig.cap = paste0("Emissions (kt) of ",params$species," month-1 sector-1")}
#| fig-pos: "h"
knitr::include_graphics(paste0("/gws/nopw/j04/ceh_generic/samtom/EMEP_inputs/outputs/EMEP4UK",params$emep_version,"/inv",params$naei_inv,"/UKEIRE/",params$tp_scheme,"/TP",tp_scheme,"_",uk_agg_schema,"/plots/e",params$y,"/",params$species,"_UKEIRE_",params$y,"emis_",params$map_yr_uk,"map_",params$naei_inv,"inv_5_UKMONSECLINE.png"))
```

N.B. as UK data starts in SNAP sectors & Eire data in GNFR sectors, UK & Eire sectoral monthly emissions are sometimes flat

```{r, message = FALSE, echo = FALSE}

dt <- copy(params$dt_month_uk)
dt[, Area := toupper(Area)]
dt[, emis_kt := round(emis_t/1000, 2)]
dt[, time := paste0("M", time)]
dt[, time := factor(time, levels = paste0("M",1:12))]
setnames(dt, "sec_name", "Sector")
dt_wide <- dcast(dt, Sector+Area~time, value.var = "emis_kt")
dt_wide[, Total := rowSums(.SD, na.rm=T), .SDcols = paste0("M",1:12)]
dt_wide <- adorn_totals(dt_wide)

kable(dt_wide[Area != "SEA"], align = "rcccccccccccccc", caption = paste0("Monthly ", params$species, " emissions (kt) per sector per ISO"), booktabs = TRUE)
 #%>%
 # kable_styling(font_size = 7)
```

\newpage

# EU INPUTS FILES

## METADATA

**Project:** NFC

**Git Repo:** https://github.com/oSamTo/EMEP_inputs

**JASMIN dir:** /gws/nopw/j04/ceh_generic/samtom/EMEP_inputs

**Output dir:** `r gsub("/gws/nopw/j04/ceh_generic/samtom/EMEP_inputs/", "", params$eu_folname)`

**EMEP version:** `r params$emep_version`

\vspace{\baselineskip}

Input files processed to 0.1$^\circ$ x 0.1$^\circ$ resolution, in WGS84. 

\vspace{\baselineskip}

```{r, message = FALSE, echo = FALSE}

dt <- data.table(Area = c("EU ISOs"),
                 Species = params$species,
                 Inv_Year = params$emep_inv,
				 Emis_Year = params$y,
				 temp_res = params$time_dim,
				 orig_sec = "GNFR",
				 orig_res = 0.1,
				 orig_crs = "WGS84")

kable(dt, align = "cccccccc", caption = "Metadata of inputs", 
	  booktabs = TRUE)
		 
```

**EMEP ISO codes:** `r paste(dt_iso[, ISO_char])`

Notes: 

* Emissions from the EU (and wider) domain remain in GNFR sectors (see 
/data/lookups/EMEP_sectors.csv) with UK+10km data masked out. EMEP/CEIP 
emissions take precedence outside of this mask. 

* Emissions data are provided in WGS84 lat lon, at 0.1$^\circ$ x 0.1$^\circ$ resolution. 

  * This is done in /gws/nopw/j04/ceh_generic/inventory_processor

\newpage

## MAP: TOTAL EMISSION ANNUM^-1^


```{r, eval=TRUE, fig.align="center", out.width = "0.9\\linewidth", echo = FALSE, fig.cap = paste0("EU total ",params$species," emissions annum-1 (UK/Eire emissions for context).")}
#| fig-pos: "h"
knitr::include_graphics(paste0("/gws/nopw/j04/ceh_generic/samtom/EMEP_inputs/outputs/EMEP4UK",params$emep_version,"/inv",params$emep_inv,"/EU/",params$tp_scheme,"/TP",tp_scheme,"_",eu_agg_schema,"/plots/e",params$y,"/",params$species,"_EU_",params$y,"emis_",params$y,"map_",params$emep_inv,"inv_8_EUTOTMAP.png"))
```

\vspace{\baselineskip}

```{r, message = FALSE, echo = FALSE, label = tab-totEU}

l_ann_tot <- params$l_eu_maps$ann_tot_all
s_ann_tot <- rast(l_ann_tot)
r_ann_tot <- app(s_ann_tot, sum, na.rm=T)
r_ann_tot[r_ann_tot == 0] <- NA

dt_tot_emis <- fread(params$l_fname_eu_sums[["ncoutput"]])

dt1 <- data.table(Area = c("EU"),
                  Species = params$species,
 				  Total_kt = dt_tot_emis[, round(sum(emis_t_tot_ncoutput, na.rm = T)/1000, 2)],
				  mean_t   = round(global(r_ann_tot, mean, na.rm=T)[,1], 2),
				  max_t    = round(global(r_ann_tot, max, na.rm=T)[,1], 2))
							  
#dt2 <- data.table(Area = c("Overall"),
#                  Species = params$species,
# 				  Total_kt = dt1[, sum(Total_kt, na.rm=T)],
#				  mean_t   = round(global(app(s_ann_tot, sum, na.rm=T), mean, na.rm=T)[,1], 2),
#				  max_t    = round(global(app(s_ann_tot, sum, na.rm=T), max, na.rm=T)[,1], 2))

#dt <- rbindlist(list(dt1, dt2), use.names = T)

kable(dt1, align = "ccccc", caption = paste0("Annual emissions of ", params$species, " per ISO."), 
	  booktabs = TRUE)
```

* N.B. EU totals in table above have UK & Eire masked out.

\newpage

## PROCESSING FLOW

Figure \ref{fig:fig-procEU1} shows total EMEP emissions as taken 
from EMEP/CEIP (inventory_data) and emissions post processing (processed_data). 

```{r, eval=TRUE, fig.align="center", out.width = "0.35\\linewidth", echo = FALSE, fig.cap = "Emissions processing from source.", label = fig-procEU1}
#| fig-pos: "h"
knitr::include_graphics(paste0("/gws/nopw/j04/ceh_generic/samtom/EMEP_inputs/outputs/EMEP4UK",params$emep_version,"/inv",params$emep_inv,"/EU/",params$tp_scheme,"/TP",tp_scheme,"_",eu_agg_schema,"/plots/e",params$y,"/",params$species,"_EU_",params$y,"emis_",params$y,"map_",params$emep_inv,"inv_9_EUINVBAR.png"))
```

Total `r params$species` lost in EU domain, following processing: `r round(fread(params$l_fname_eu_sums[["inventory"]])[,sum(emis_t, na.rm=T)/1000]- fread(params$l_fname_eu_sums[["processed"]])[,sum(ann_emis_kt, na.rm=T)], 1)` kt. Spatial breakdown :

```{r, message = FALSE, echo = FALSE, label = EU-data-lost}

dt_inv <- fread(params$l_fname_eu_sums[["inventory"]])
dt_inv <- dt_inv[, list(inventory_kt = round(sum(emis_t, na.rm=T)/1000, 1)), by = .(ISO)]
							  
dt_pro <- fread(params$l_fname_eu_sums[["processed"]])
dt_pro <- dt_pro[, list(processed_kt = round(sum(ann_emis_kt, na.rm=T), 1)), by = .(ISO)]

dt <- dt_inv[dt_pro, on = "ISO"]
dt[, lost_kt := round(processed_kt - inventory_kt, 2)]
dt[, proc_ratio := round(processed_kt / inventory_kt, 3)]

dt <- dt[proc_ratio < 1]
dt <- adorn_totals(dt)

kable(dt, align = "ccccc", caption = paste0("Emissions lost in EMEP domain per ISO."), 
	  booktabs = TRUE)
```


Figure \ref{fig:fig-procEU2} shows the data progressing to NetCDF inputs - emissions inside 10km sea buffer are masked and not used for EMEP-EU.

```{r, eval=TRUE, fig.align="center", out.width = "0.6\\linewidth", echo = FALSE, fig.cap = "Processed toals to NetCDF input data.", label = fig-procEU2}
#| fig-pos: "h"
knitr::include_graphics(paste0("/gws/nopw/j04/ceh_generic/samtom/EMEP_inputs/outputs/EMEP4UK",params$emep_version,"/inv",params$emep_inv,"/EU/",params$tp_scheme,"/TP",tp_scheme,"_",eu_agg_schema,"/plots/e",params$y,"/",params$species,"_EU_",params$y,"emis_",params$y,"map_",params$emep_inv,"inv_10_EUNCBAR.png"))
```

\newpage

## EMISSIONS SECTORAL CLASSIFICATIONS

### EU data starts in GNFR sectors. 

```{r, message = FALSE, echo = FALSE}

dt <- params$dt_sec[,-1]

kable(dt, align = "rcccccl", caption = "Sectoral lookup table, SNAP <--> GNFR", 
	  booktabs = TRUE)
```

\newpage

## MAP: SECTORAL EMISSIONS ANNUM^-1^

### EU temporal schema: `r params$time_dim`

Source: `r if(params$time_dim == "annual"){paste0("EMEP ", params$emep_version, " schema: data/temporal/EMEP4UK",params$emep_version,"/")}else{paste0("UKEM ", params$tp_scheme, " schema: /gws/nopw/j04/ukem/test/sam/ukem_pro/output/")}`

```{r, eval=TRUE, fig.align="center", out.width = "0.85\\linewidth", echo = FALSE, fig.cap = paste0("EMEP-EU total ",params$species," emissions sector-1 annum-1")}
#| fig-pos: "h"
knitr::include_graphics(paste0("/gws/nopw/j04/ceh_generic/samtom/EMEP_inputs/outputs/EMEP4UK",params$emep_version,"/inv",params$emep_inv,"/EU/",params$tp_scheme,"/TP",tp_scheme,"_",eu_agg_schema,"/plots/e",params$y,"/",params$species,"_EU_",params$y,"emis_",params$y,"map_",params$emep_inv,"inv_13_EUANNSECMAP.png"))
```

```{r, message = FALSE, echo = FALSE}

dt_out <- fread(params$l_fname_eu_sums[["ncoutput"]])
setnames(dt_out, "iso_char", "ISO")

dt_long_eu <- dt_out[,list(emis_kt = round(sum(emis_t_tot_ncoutput, na.rm=T)/1000,2)), by = .(sec_name)]
dt_long_eu[, ISO := "EU"]
dt_long_iso <- dt_out[ISO %in% c("FR","NL","BE","DE","ES","NOS","ATL"), list(emis_kt = round(sum(emis_t_tot_ncoutput, na.rm=T)/1000,2)), by = .(ISO, sec_name)]
dt_long <- rbindlist(list(dt_long_eu, dt_long_iso), use.names = T)
dt_long[, ISO := factor(ISO, levels = c("EU","FR","NL","BE","DE","ES","NOS","ATL"))]

dt_wide <- dcast(dt_long, sec_name~ISO, value.var = "emis_kt")
# dt_wide[, Total := rowSums(.SD, na.rm=T), .SDcols = setdiff(names(dt_wide), "ISO")]
dt_wide <- adorn_totals(dt_wide)
setnames(dt_wide, "sec_name", "Sector")

kable(dt_wide, align = "cc", caption = paste0(params$species," emissions (kt) per sector in EU, plus UK proximal ISOs"), 
	  booktabs = TRUE)
```

\newpage

## EMISSIONS MONTH^-1^

### EMEP-EU temporal schema: `r params$time_dim`

Source: `r if(params$time_dim == "annual"){paste0("EMEP ", params$emep_version, " schema: data/temporal/EMEP4UK",params$emep_version,"/")}else{paste0("UKEM ", params$tp_scheme, " schema: /gws/nopw/j04/ukem/test/sam/ukem_pro/output/")}`

```{r, eval=TRUE, fig.align="center", out.width = "0.6\\linewidth", echo = FALSE, fig.cap = paste0("Emissions of ",params$species," month-1"), label = "fig-montotEU"}
#| fig-pos: "h"
knitr::include_graphics(paste0("/gws/nopw/j04/ceh_generic/samtom/EMEP_inputs/outputs/EMEP4UK",params$emep_version,"/inv",params$emep_inv,"/EU/",params$tp_scheme,"/TP",tp_scheme,"_",eu_agg_schema,"/plots/e",params$y,"/",params$species,"_EU_",params$y,"emis_",params$y,"map_",params$emep_inv,"inv_11_EUMONTOTLINE.png"))
```

Total monthly emissions of `r params$species` in EU and ISOs nearest to UK/Eire. 
`r if(params$time_dim == "annual"){paste0("As annual emissions input into NetCDF, using model profiles.")}`
`r if(params$time_dim == "month"){paste0("Temporal profiles taken from UKEM")}`

```{r, message = FALSE, echo = FALSE}

dt_mon_eu <- params$dt_month_eu[,list(emis_kt = round(sum(emis_t, na.rm=T)/1000, 1)), by = .(time)]
dt_mon_eu[, ISO := "EU"]
dt_mon_iso <- params$dt_month_eu[ISO %in% c("FR","NL","BE","DE","ES","NOS","ATL"),list(emis_kt = round(sum(emis_t, na.rm=T)/1000, 1)), by = .(ISO, time)]
dt <- rbindlist(list(dt_mon_eu, dt_mon_iso), use.names = T)

dt[, time := gsub("t", "M", time)]
dt[, time := factor(time, levels = paste0("M",1:12))]
dt[, ISO := factor(ISO, levels = c("EU","FR","NL","BE","DE","ES","NOS","ATL"))]

dt_wide <- dcast(dt, ISO~time, value.var = "emis_kt")
dt_wide[, Total := rowSums(.SD, na.rm=T), .SDcols = paste0("M",1:12)]

kable(dt_wide, align = "rccccccccccccc", caption = paste0("Total monthly ", params$species, " emissions (kt)"), 
	  booktabs = TRUE)
```

\newpage

## SECTORAL EMISSIONS MONTH^-1^

### EMEP-EU temporal schema: `r params$time_dim`

Source: `r if(params$time_dim == "annual"){paste0("EMEP ", params$emep_version, " schema: data/temporal/EMEP4UK",params$emep_version,"/")}else{paste0("UKEM ", params$tp_scheme, " schema: /gws/nopw/j04/ukem/test/sam/ukem_pro/output/")}`

```{r, eval=TRUE, fig.align="center", out.width = "0.7\\linewidth", echo = FALSE, fig.cap = paste0("Emissions (kt) of ",params$species," month-1 sector-1")}
#| fig-pos: "h"
knitr::include_graphics(paste0("/gws/nopw/j04/ceh_generic/samtom/EMEP_inputs/outputs/EMEP4UK",params$emep_version,"/inv",params$emep_inv,"/EU/",params$tp_scheme,"/TP",tp_scheme,"_",eu_agg_schema,"/plots/e",params$y,"/",params$species,"_EU_",params$y,"emis_",params$y,"map_",params$emep_inv,"inv_12_EUMONSECLINE.png"))
```

```{r, message = FALSE, echo = FALSE}

dt <- copy(params$dt_month_eu)
dt <- dt[, list(emis_kt = round(sum(emis_t, na.rm=T)/1000, 1)), by = .(sec_name, time)]
dt[, ISO := "EU"]

dt[, time := gsub("t", "M", time)]
dt[, time := factor(time, levels = paste0("M",1:12))]
setnames(dt, "sec_name", "Sector")
dt_wide <- dcast(dt, Sector+ISO~time, value.var = "emis_kt")
dt_wide[, Total := rowSums(.SD, na.rm=T), .SDcols = paste0("M",1:12)]
dt_wide <- adorn_totals(dt_wide)

kable(dt_wide, align = "rcccccccccccccc", caption = paste0("Monthly ", params$species, " emissions (kt) per sector per ISO"), 
	  booktabs = TRUE)
```

\newpage

# TIME SERIES