---
params:
  project: NA
  scenario: NA
  y: NA
  species: NA
  domain: NA
  folname: NA
  inv: NA
  data_source: NA
  map_yr_uk: NA
  time_dim: NA
  emep_version: NA
  v_EMEP_sec: NA
  agg_schema: NA
  tp_scheme: NA
  dt_sec: NA
  dt_poll: NA
  summary_paths: NA
  raster_paths: NA
  plot_paths: NA
title: "EMEP input files QAQC"
author: "Sam Tomlinson"
date: "`r Sys.Date()`"
header-includes:
 \usepackage{geometry}
 \geometry{top=1cm,left=1cm,bottom=1cm,right=1cm}
output:
  pdf_document:
    fig_caption: true
    extra_dependencies: ["booktabs"]
---

```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
library(data.table)
library(knitr)
library(kableExtra)
library(janitor)
library(terra)
root <- "/gws/ssde/j25b/ceh_generic/samtom/EMEP_inputs"
source(file.path(root, "R/qaqc_codex_v2/utils.R"))
source(file.path(root, "R/qaqc_codex_v2/data.R"))
source(file.path(root, "R/qaqc_codex_v2/plots.R"))

kbl <- function(...) {
  print(knitr::kable(...))
  invisible(NULL)
}

include_plot <- function(path) {
  path <- as.character(path)[1]
  if (!is.na(path) && file.exists(path)) {
    knitr::include_graphics(path, rel_path = FALSE)
  } else {
    cat("Plot not available.")
  }
}

domain_label <- switch(params$domain, UKEIRE = "UK/Eire", EU = "EU", GLOBAL = "Global", params$domain)
summary_paths <- params$summary_paths
raster_paths <- params$raster_paths
plot_paths <- params$plot_paths
```

# `r domain_label` Input Files

## Metadata

**Project:** `r params$project`

**Scenario:** `r params$scenario`

**Domain:** `r params$domain`

**Species:** `r params$species`

**Emissions year:** `r params$y`

**Inventory / source version:** `r params$inv`

**Data source:** `r params$data_source`

**Output dir:** `r gsub("/gws/ssde/j25b/ceh_generic/samtom/EMEP_inputs/", "", params$folname)`

**EMEP version:** `r params$emep_version`

```{r metadata-table, echo = FALSE, results = "asis"}

eire_sec <- if(params$species == "hcl")	{
	          "zhang_2022"
	        } else {
		      "GNFR"
		    }

eire_res <- if(params$species == "hcl")	{
	          "0.1_deg"
	        } else {
		      "1km"
		    }

eire_crs <- if(params$species == "hcl")	{
	          "LL"
	        } else {
		      "TM65"
		    }

if(params$domain == "UKEIRE"){

  dt_meta <- data.table(
    Domain = params$domain,
    Species = params$species,
    Area = c("UK", "Eire", "SEA"),
    Emis_Year = params$y,
    Source = c(params$data_source,"EMEP", "N/A"),
    Inv = params$inv,
    orig_sec = c("SNAP", eire_sec, "N/A"),
		orig_res = c("1km", eire_res, "N/A"),
    Temporal = params$time_dim    
  )
  kbl(dt_meta, align = "ccccccccc", caption = "Run metadata", booktabs = TRUE)

  }else{

  dt_meta <- data.table(
    Domain = params$domain,
    Species = params$species,
    Emis_Year = params$y,
    Source = params$data_source,
    Inv = params$inv,
    Temporal = params$time_dim,
    Aggregation = params$agg_schema
  )
  kbl(dt_meta, align = "ccccccc", caption = "Run metadata", booktabs = TRUE)
}

```

```{r, results='asis', echo=FALSE}
if (params$domain == "UKEIRE") {

cat("
Notes: 

* Emissions from the UK and Eire are processed to consistent sector categories (see 
/data/lookups/EMEP_sectors.csv) and masked to a UK+10km mask. EMEP/CEIP 
emissions take precedence outside of this mask. 

* The 10km strip from the land territory into the sea is combined (UK + Eire) to a 'SEA' ISO 
code, for all sectors.

* Emissions maps are pre-processed from 1km x 1km BNG (UK) and 1km x 1km Irish 
Grid TM65 (Eire) into WGS84 lat lon, at 0.01$^\\circ$ x 0.01$^\\circ$ resolution. 

  * HCl data for Eire/EU is taken from Zhang et al. (2022) - https://doi.org/10.5194/essd-14-5297-2022

  * This is done in /gws/ssde/j25b/ceh_generic/inventory_processor

* Totals are adherent to NFR level reporting (i.e. maps are scaled)

  * Taken from: https://naei.energysecurity.gov.uk/data/data-selector
")

}else if(params$domain == "EU"){

cat("
Notes: 

* Emissions from the EU (and wider) domain remain in GNFR sectors (see 
/data/lookups/EMEP_sectors.csv) with UK+10km data masked out. EMEP/CEIP 
emissions take precedence outside of this mask. 

* Emissions data are provided in WGS84 lat lon, at 0.1$^\\circ$ x 0.1$^\\circ$ resolution. 

  * This is done in /gws/ssde/j25b/ceh_generic/inventory_processor

")

}
```


\newpage

## Processing Flow

```{r stage-plot, echo = FALSE, fig.align = "center", out.width = "0.95\\linewidth", fig.cap = "Processing-stage totals by ISO, normalised to each ISO inventory total. Inventory is therefore 1."}
include_plot(plot_paths$stage)
```

```{r stage-table, echo = FALSE, results = "asis"}
dt_stage <- qaqc_v2_stage_totals(summary_paths)
if (nrow(dt_stage) > 0) {
  dt <- dt_stage[, .(Total_kt = round(sum(emis_t, na.rm = TRUE) / 1000, 3)), by = .(Area, stage)]
  dt_domain <- dt[, .(Total_kt = sum(Total_kt, na.rm = TRUE)), by = stage]
  ref <- dt_domain[stage == "inventory", Total_kt][1]
  if (!is.na(ref)) {
    dt_domain[, ratio_to_inventory := round(Total_kt / ref, 6)]
    dt_domain[, delta_from_inventory_kt := round(Total_kt - ref, 6)]
  }
  kbl(dt_domain, caption = "Domain-level conservation check", booktabs = TRUE)
} else {
  kbl(data.table(Note = "No processing-stage summaries found."), booktabs = TRUE)
}
```

```{r inventory-processed-sector-loss, echo = FALSE, results = "asis"}
dt_sector_loss <- qaqc_v2_inventory_processed_sector_loss(summary_paths, params$dt_sec)
if (nrow(dt_sector_loss) > 0) {
  kbl(
    dt_sector_loss,
    caption = paste0(
      "GNFR sectors contributing to inventory-to-processed ",
      params$species,
      " losses (kt). Missing_pct is percent of total positive loss."
    ),
    booktabs = TRUE
  )
} else {
  kbl(
    data.table(Note = "No positive inventory-to-processed sector losses found, or sector columns were unavailable."),
    booktabs = TRUE
  )
}
```

\newpage

## Total Emissions Map

Map classes use quantiles with probabilities `0.20, 0.40, 0.55, 0.70, 0.82, 0.92, 0.97, 1.00` and the `Spectral` brewer palette. For the global domain, ISO polygons are drawn on this total map only.

```{r total-map, echo = FALSE, fig.align = "center", out.width = "0.95\\linewidth", fig.cap = "Total emissions from precomputed QAQC raster (data inside .nc file)."}
include_plot(plot_paths$raster_total)
```

```{r raster-summary, echo = FALSE, results = "asis"}
dt_rast <- qaqc_v2_raster_summary(raster_paths)
if (nrow(dt_rast) > 0) {
  kbl(dt_rast, caption = "Precomputed QAQC raster summaries (data inside .nc input file)", booktabs = TRUE)
} else {
  kbl(data.table(Note = "No precomputed QAQC rasters found."), booktabs = TRUE)
}
```

\newpage

## Sectoral Emissions Maps

```{r sector-map, echo = FALSE, fig.align = "center", out.width = "0.98\\linewidth", fig.cap = "Sectoral emissions from precomputed QAQC rasters."}
include_plot(plot_paths$raster_sector)
```

\newpage

## Sectoral Emissions: Top ISO Codes

Figure 4 and Table 5 are restricted to the ten ISO codes with the biggest totals across all sectors.

```{r sector-plot, echo = FALSE, fig.align = "center", out.width = "0.82\\linewidth", fig.cap = "Sectoral emissions for the ten largest ISO totals."}
include_plot(plot_paths$sector_bar)
```

```{r sector-table, echo = FALSE, results = "asis"}
dt_sector <- qaqc_v2_top_iso_sector_table(summary_paths, top_n = 10)
if (nrow(dt_sector) > 0) {
  if(params$domain == "UKEIRE"){
    dt_sector[, ncTotal := rowSums(.SD, na.rm=T), .SDcols = c("GB", "IE", "SEA")]
  }
  kbl(dt_sector, caption = paste0(params$species, " emissions by sector for top ten ISO totals (kt). ISO columns are ordered largest to smallest."), booktabs = TRUE)
} else {
  kbl(data.table(Note = "No sectoral NetCDF output summary found."), booktabs = TRUE)
}
```

\newpage

```{r global-timeseries-title, echo = FALSE, results = "asis"}
if (params$domain == "GLOBAL") {
  cat("## Global Inventory Time Series\n\n")
} else {
  cat("## Inventory Time Series\n\n")
}
```

```{r ts-plots, echo = FALSE, fig.align = "center", out.width = "0.82\\linewidth", fig.cap = "Global inventory total over time."}
if (params$domain == "GLOBAL") {
  include_plot(plot_paths$timeseries_total)
}
```

```{r ts-sector-plot, echo = FALSE, fig.align = "center", out.width = "0.9\\linewidth", fig.cap = "Global inventory sector totals over time."}
if (params$domain == "GLOBAL") {
  include_plot(plot_paths$timeseries_sector)
}
```

```{r ts-table, echo = FALSE, results = "asis"}
if (params$domain == "GLOBAL") {
  ts <- qaqc_v2_global_timeseries(params$data_source, params$inv, params$species, params$y)
  if (nrow(ts$data) > 0) {
    dt <- copy(ts$data)
    dt[, emis_kt := round(emis_kt, 3)]
    kbl(dt, caption = paste0("Global ", params$species, " inventory totals from ", basename(ts$file), " (kt)"), booktabs = TRUE)
  } else {
    kbl(data.table(Note = "No global time-series data found."), booktabs = TRUE)
  }
}
```

\newpage

## Monthly Emissions

```{r monthly-table, echo = FALSE, results = "asis"}
dt_out <- qaqc_v2_read_csv(summary_paths$ncoutput)
time_cols <- grep("^t[0-9]+$", names(dt_out), value = TRUE)
if (nrow(dt_out) > 0 && length(time_cols) > 1) {
  area_col <- intersect(c("Area", "ISO", "iso_char"), names(dt_out))[1]
  if (!is.na(area_col)) data.table::setnames(dt_out, area_col, "Area")
  id_cols <- setdiff(names(dt_out), time_cols)
  dt <- melt(dt_out, id.vars = id_cols, measure.vars = time_cols, variable.name = "time", value.name = "emis_t")
  dt <- dt[, .(emis_kt = round(sum(emis_t, na.rm = TRUE) / 1000, 3)), by = .(Area, time)]
  dt_wide <- dcast(dt, Area ~ time, value.var = "emis_kt", fill = 0)
  kbl(dt_wide, caption = paste0("Monthly ", params$species, " NetCDF output emissions (kt)"), booktabs = TRUE)
} else {
  kbl(data.table(Note = "No monthly columns available. Annual-only NetCDF summaries contain t1 only."), booktabs = TRUE)
}
```

\newpage

## Sector Classifications

```{r sector-lookup, echo = FALSE, results = "asis"}
if (!is.null(params$dt_sec) && nrow(params$dt_sec) > 0) {
  dt <- as.data.table(params$dt_sec)
  dt[, names(dt) := lapply(.SD, function(x) if (is.character(x)) iconv(x, from = "", to = "UTF-8", sub = "") else x)]
  kbl(dt, caption = "Sectoral lookup table", booktabs = TRUE)
} else {
  kbl(data.table(Note = "No sector lookup table supplied."), booktabs = TRUE)
}
```
