---
params:
  project: NA
  scenario: NA
  y: NA
  species: NA
  domain: NA
  folname: NA
  inv: NA
  data_source: NA
  map_yr: NA
  time_dim: NA
  emep_version: NA
  v_EMEP_sec: NA
  agg_schema: NA
  tp_scheme: NA
  dt_sec: NA
  dt_poll: NA
  summary_paths: NA
  raster_paths: NA
  plot_paths: NA
  v_snap_split: NA
  global_iso: NA
title: "EMEP input files QAQC"
author: "Sam Tomlinson"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    toc: false
    toc_float: false
    number_sections: false
    theme: flatly
    df_print: paged
    code_folding: hide
    self_contained: true
    fig_caption: true
---

```{r setup, include=FALSE}
# HTML-friendly QAQC report helpers (paths, plot embedding, table helpers)
project_root <- knitr::opts_knit$get("root.dir")
if (is.null(project_root) || is.na(project_root) || !nzchar(project_root)) {
  project_root <- getwd()
}
project_root <- normalizePath(project_root, mustWork = FALSE)

knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center",
  out.width = "100%"
)

suppressPackageStartupMessages({
  library(data.table)
  library(knitr)
  library(kableExtra)
  library(janitor)
  library(terra)
})

`%||%` <- function(x, y) {
  if (is.null(x) || length(x) == 0 || all(is.na(x))) y else x
}

q_is_abs_path <- function(path) {
  grepl("^([A-Za-z]:[\\\\/]|/|\\\\\\\\)", path)
}

q_norm_slash <- function(path) {
  gsub("\\\\", "/", path)
}

q_abs_path <- function(path) {
  if (is.null(path) || length(path) == 0 || all(is.na(path))) {
    return(NA_character_)
  }

  path <- as.character(path[[1]])
  path <- q_norm_slash(path)

  if (q_is_abs_path(path)) {
    return(normalizePath(path, winslash = "/", mustWork = FALSE))
  }

  candidates <- c(
    file.path(project_root, path),
    path
  )

  existing <- candidates[file.exists(candidates)]
  if (length(existing) > 0) {
    return(normalizePath(existing[[1]], winslash = "/", mustWork = FALSE))
  }

  normalizePath(file.path(project_root, path), winslash = "/", mustWork = FALSE)
}

# report_dir is expected to be the parent of the plots folder
plot_dir_abs <- function(plot_dir) {
  q_abs_path(plot_dir)
}

q_plot_rel <- function(path, plot_dir) {
  if (is.null(path) || length(path) == 0 || all(is.na(path))) return(NA_character_)

  vapply(as.character(path), function(p) {
    p <- q_norm_slash(p)
    report_dir <- normalizePath(file.path(plot_dir, ".."), winslash = "/", mustWork = FALSE)
    expected <- file.path(report_dir, "plots", basename(p))

    if (file.exists(expected)) {
      return(q_norm_slash(file.path("plots", basename(p))))
    }

    p_abs <- q_abs_path(p)
    if (file.exists(p_abs) && requireNamespace("xfun", quietly = TRUE)) {
      rel <- xfun::relative_path(
        normalizePath(p_abs, winslash = "/", mustWork = TRUE),
        normalizePath(report_dir, winslash = "/", mustWork = FALSE)
      )
      return(q_norm_slash(rel))
    }

    q_norm_slash(file.path("plots", basename(p)))
  }, character(1))
}

q_html_escape <- function(x) {
  x <- as.character(x)
  x[is.na(x)] <- ""
  x <- gsub("&", "&amp;", x, fixed = TRUE)
  x <- gsub("<", "&lt;", x, fixed = TRUE)
  x <- gsub(">", "&gt;", x, fixed = TRUE)
  x <- gsub('"', "&quot;", x, fixed = TRUE)
  x
}

q_include_plot <- function(path, plot_dir = NULL, width = NULL, caption = NULL, max_width = NULL) {
  paths <- as.character(unlist(path, use.names = FALSE))
  paths <- paths[!is.na(paths) & nzchar(paths)]

  if (length(paths) == 0) {
    return(knitr::asis_output("<p><em>No image path was supplied.</em></p>"))
  }

  # Pick up chunk options when not supplied directly
  width <- width %||% knitr::opts_current$get("out.width") %||% "100%"
  caption <- caption %||% knitr::opts_current$get("fig.cap")

  report_dir <- if (is.character(plot_dir) && length(plot_dir) == 1 && !is.na(plot_dir)) {
    normalizePath(plot_dir, winslash = "/", mustWork = FALSE)
  } else {
    project_root
  }

  # Optional cap in pixels, useful because 90% of a 1500px container is still huge
  max_width_css <- if (!is.null(max_width)) {
    paste0(" max-width:", max_width, ";")
  } else {
    ""
  }

  blocks <- vapply(seq_along(paths), function(i) {
    p <- paths[[i]]
    full <- q_abs_path(p)

    if (!file.exists(full)) {
      full2 <- file.path(report_dir, "plots", basename(p))
      if (file.exists(full2)) {
        full <- normalizePath(full2, winslash = "/", mustWork = TRUE)
      }
    }

    if (!file.exists(full)) {
      return(paste0(
        "<p class=\"missing-plot\"><strong>Missing plot:</strong> ",
        q_html_escape(p),
        "</p>"
      ))
    }

    img_src <- knitr::image_uri(full)

    this_caption <- if (!is.null(caption) && length(caption) >= i && !is.na(caption[[min(i, length(caption))]])) {
      caption[[min(i, length(caption))]]
    } else {
      basename(full)
    }

    paste0(
      "<figure class=\"plot-figure\" style=\"width:", q_html_escape(width),
      ";", max_width_css, " margin-left:auto; margin-right:auto;\">",
      "<img class=\"report-plot\" src=\"", q_html_escape(img_src),
      "\" alt=\"", q_html_escape(basename(full)), "\">",
      "<figcaption>", q_html_escape(this_caption), "</figcaption>",
      "</figure>"
    )
  }, character(1))

  wrapper_class <- if (length(blocks) > 1) "plot-grid" else "plot-single"

  knitr::asis_output(paste0(
    "<div class=\"", wrapper_class, "\">",
    paste0(blocks, collapse = "\n"),
    "</div>"
  ))
}

include_plot <- function(path, width = NULL, caption = NULL, max_width = NULL) {
  q_include_plot(path, width = width, caption = caption, max_width = max_width)
}

# HTML-friendly table helper using kableExtra
kbl <- function(
  dt,
  caption = NULL,
  scroll = FALSE,
  max_height = "600px",
  max_width = "100%",
  ...
) {
  if (is.data.frame(dt) || data.table::is.data.table(dt)) {

    # Manual caption, so captions are independent of scroll_box behaviour
    if (!is.null(caption)) {
      cat(
        paste0(
          "<div class=\"table-caption\">",
          q_html_escape(caption),
          "</div>\n"
        )
      )
    }

    tbl <- knitr::kable(
      dt,
      format = "html",
      escape = FALSE,
      ...
    )

    if (requireNamespace("kableExtra", quietly = TRUE)) {
      tbl <- kableExtra::kable_styling(
        tbl,
        bootstrap_options = c("striped", "hover", "condensed"),
        full_width = FALSE
      )
    }

    # Only add scrollbars where explicitly requested
    if (scroll && requireNamespace("kableExtra", quietly = TRUE)) {
      tbl <- kableExtra::scroll_box(
        tbl,
        width = max_width,
        height = max_height
      )
    }

    print(tbl)

  } else {
    if (!is.null(caption)) {
      cat(
        paste0(
          "<div class=\"table-caption\">",
          q_html_escape(caption),
          "</div>\n"
        )
      )
    }

    print(knitr::kable(dt, format = "html", escape = FALSE, ...))
  }

  invisible(NULL)
}

dtbl <- function(
  dt,
  caption = NULL,
  scroll_y = "550px"
) {
  if (!requireNamespace("DT", quietly = TRUE)) {
    stop("Package 'DT' is required for interactive tables.")
  }

  if (!is.null(caption)) {
    cat(
      paste0(
        "<div class=\"table-caption\">",
        q_html_escape(caption),
        "</div>\n"
      )
    )
  }

  DT::datatable(
    dt,
    rownames = FALSE,
    filter = "top",
    class = "compact stripe hover",
    options = list(
      paging = FALSE,
      scrollX = TRUE,
      scrollY = scroll_y,
      scrollCollapse = TRUE,
      autoWidth = TRUE,
      order = list(list(0, "asc")),
      dom = "frt"
    )
  )
}

root <- "/gws/ssde/j25b/ceh_generic/samtom/EMEP_inputs"
source(file.path(root, "R/qaqc_codex_v2/utils.R"))
source(file.path(root, "R/qaqc_codex_v2/data.R"))
source(file.path(root, "R/qaqc_codex_v2/plots.R"))

# expose params-derived helpers
summary_paths <- params$summary_paths
raster_paths <- params$raster_paths
plot_paths <- params$plot_paths
```

```{css, echo=FALSE}
body { font-size: 15px; line-height: 1.5; }
.main-container { max-width: 1500px !important; }
h1, h2, h3 { font-weight: 750; }
.table {
  width: auto !important;
  margin-left: 0;
  margin-right: auto;
}
.table-caption {
  caption-side: top;  
  font-weight: 700;
  color: #1f2933;
  margin: 0.75rem 0 0.25rem 0;
}
img { 
  max-width: 100%; 
  height: auto; 
  border: 1px solid #d9e2ec; 
  border-radius: 0.65rem; 
  box-shadow: 0 1px 5px rgba(0,0,0,0.08); 
  background: white; 
  margin-bottom: 1.1rem; }
.table-scroll {
  overflow: auto;
  width: 100%;
  border: 1px solid #d9e2ec;
  border-radius: 0.45rem;
  margin: 0.75rem 0 1.4rem 0;
}
.report-table { margin-bottom: 0 !important; white-space: nowrap; }
.plot-single, .plot-grid { margin: 1rem 0 1.6rem 0; }
.plot-grid { display: grid; grid-template-columns: repeat(auto-fit, minmax(420px, 1fr)); gap: 1rem; align-items: start; }
.plot-figure {
  padding: 0.75rem;
  border: 1px solid #d9e2ec;
  border-radius: 0.5rem;
  background: #fff;
  box-sizing: border-box;
}
.report-plot {
  display: block;
  width: 100%;
  height: auto;
  margin-bottom: 0.4rem;
}
.plot-figure figcaption { margin-top: 0.45rem; font-size: 0.85rem; color: #52606d; word-break: break-all; }
.missing-plot { color: #9b1c1c; background: #fff5f5; border: 1px solid #f8b4b4; border-radius: 0.4rem; padding: 0.75rem; }
```

# `r switch(params$domain, UKEIRE = "UK/Eire", EU = "EU", GLOBAL = "Global", params$domain)` Input Files

## Metadata

**Project:** `r params$project`

**Scenario:** `r params$scenario`

**Domain:** `r params$domain`

**SNAPs split (GB):** `r params$v_snap_split`

**Output dir:** `r gsub("/gws/ssde/j25b/ceh_generic/samtom/EMEP_inputs/", "", params$folname)`

```{r metadata-table, echo = FALSE, results = "asis"}
# preserve original metadata rendering but using HTML kbl

if(params$domain == "UKEIRE"){
  eire_sec <- if(params$species == "hcl"){"zhang_2022"} else {"GNFR"}
  eire_res <- if(params$species == "hcl"){"0.1_deg"} else {"1km"}
  eire_crs <- if(params$species == "hcl"){"LL"} else {"TM65"}

  dt_meta <- data.table(
    Domain = params$domain,
    Species = params$species,
    Area = c("UK", "Eire", "SEA"),
    Emis_Year = params$y,
    Map_Year = c(params$map_yr, "2019", "N/A"),
    Source = c(params$data_source,"EMEP", "N/A"),
    Inv = params$inv,
    orig_sec = c("SNAP", eire_sec, "N/A"),
    orig_res = c("1km", eire_res, "N/A"),
    Temporal = params$time_dim,
    EMEP_version = params$emep_version
  )
  kbl(dt_meta, caption = "Run metadata")

} else {
  dt_meta <- data.table(
    Domain = params$domain,
    Species = params$species,
    Emis_Year = params$y,
    Map_Year = params$map_yr,
    Source = params$data_source,
    Inv = params$inv,
    Temporal = params$time_dim,
    Aggregation = params$agg_schema,
    ISO_coding = params$global_iso,
    EMEP_version = params$emep_version
  )
  kbl(dt_meta, caption = "Run metadata")
}
```

```{r, results='asis', echo=FALSE}
if (params$domain == "UKEIRE") {
cat("\nNotes: \n\n* Emissions from the UK and Eire are processed to consistent sector categories (see /data/lookups/EMEP_sectors.csv) and masked to a UK+10km mask. EMEP/CEIP emissions take precedence outside of this mask. \n\n* The 10km strip from the land territory into the sea is combined (UK + Eire) to a 'SEA' ISO code, for all sectors.\n\n* Emissions maps are pre-processed from 1km x 1km BNG (UK) and 1km x 1km Irish Grid TM65 (Eire) into WGS84 lat lon, at 0.01$^\\circ$ x 0.01$^\\circ$ resolution. \n\n  * HCl data (if used) for Eire/EU is taken from Zhang et al. (2022) - https://doi.org/10.5194/essd-14-5297-2022\n\n  * This is done in /gws/ssde/j25b/ceh_generic/inventory_processor\n\n* Totals are adherent to NFR level reporting (i.e. maps are scaled)\n\n  * Taken from: https://naei.energysecurity.gov.uk/data/data-selector\n")

}else if(params$domain == "EU"){
cat("\nNotes: \n\n* Emissions from the EU (and wider) domain remain in GNFR sectors (see /data/lookups/EMEP_sectors.csv) with UK+10km data masked out. EMEP/CEIP emissions take precedence outside of this mask. \n\n* Emissions data are provided in WGS84 lat lon, at 0.1$^\\circ$ x 0.1$^\\circ$ resolution. \n\n  * This is done in /gws/ssde/j25b/ceh_generic/inventory_processor\n\n")
}
```


## Processing Flow

```{r stage-plot, echo = FALSE, fig.align = "center", out.width = "60%", fig.cap = "Processing-stage totals by ISO, normalised to each ISO inventory total. Inventory is therefore 1."}
include_plot(plot_paths$stage)
```

```{r stage-table, echo = FALSE, results = "asis"}
dt_stage <- qaqc_v2_stage_totals(summary_paths, params$domain)
if (nrow(dt_stage) > 0) {
  dt <- dt_stage[, .(Total_kt = round(sum(emis_t, na.rm = TRUE) / 1000, 3)), by = .(Area, stage)]

  dt_domain <- dt[, .(Total_kt = sum(Total_kt, na.rm = TRUE)), by = stage]

  ref <- dt_domain[stage == "inventory", Total_kt][1]

  if (!is.na(ref)) {
    dt_domain[, ratio_to_inventory := round(Total_kt / ref, 6)]
    dt_domain[, delta_from_inventory_kt := round(Total_kt - ref, 6)]
  }
  kbl(dt_domain, caption = "Domain-level conservation check")
} else {
  kbl(data.table(Note = "No processing-stage summaries found."))
}
```

```{r inventory-processed-sector-loss, echo = FALSE, results = "asis"}
dt_sector_loss <- qaqc_v2_inventory_processed_sector_loss(summary_paths, params$dt_sec)
if (nrow(dt_sector_loss) > 0) {
  kbl(
    dt_sector_loss,
    caption = paste0(
      "GNFR sectors contributing to inventory-to-input ",
      params$species,
      " changes (kt). Change_pct is percent of GNFR sector. (UK + IE + SEA)"
    )
  )
} else {
  kbl(
    data.table(Note = "No inventory-to-input sector changes found, or sector columns were unavailable.")
  )
}
```

## Total Emissions Map

Map classes use quantiles with probabilities `0.20, 0.40, 0.55, 0.70, 0.82, 0.92, 0.97, 1.00` and the `Spectral` brewer palette. For the global domain, ISO polygons are drawn on this total map only.

```{r total-map, echo = FALSE, fig.align = "center", out.width = "75%", fig.cap = "Total emissions from precomputed QAQC raster (data inside .nc file)."}
include_plot(plot_paths$raster_total)
```

```{r raster-summary, echo = FALSE, results = "asis"}
dt_rast <- qaqc_v2_raster_summary(raster_paths)
if (nrow(dt_rast) > 0) {
  kbl(dt_rast, caption = "Precomputed QAQC raster summaries (data inside .nc input file)")
} else {
  kbl(data.table(Note = "No precomputed QAQC rasters found."))
}
```

## Sectoral Emissions Maps

`r if (is.null(v_snap_split)) "**NB:** GB sectors split by GNFR maps: NONE" else paste0("**NB:** GB sectors split by GNFR maps: ", paste(v_snap_split, collapse = ", "))`

```{r sector-map, echo = FALSE, fig.align = "center", out.width = "98%", fig.cap = "Sectoral emissions from precomputed QAQC rasters."}
include_plot(plot_paths$raster_sector)
```

## Sectoral Emissions: Top ISO Codes

Figure 4 and Table 5 are restricted to the ten ISO codes with the biggest totals (Gg) across all sectors.

```{r sector-plot, echo = FALSE, fig.align = "center", out.width = "70%", fig.cap = "Sectoral emissions (Gg) for the ten largest ISO totals."}
include_plot(plot_paths$sector_bar)
```

```{r sector-table, echo = FALSE, results = "asis"}
dt_sector <- qaqc_v2_top_iso_sector_table(summary_paths, top_n = 10)
if (nrow(dt_sector) > 0) {
  if(params$domain == "UKEIRE"){
    dt_sector[, ncTotal := rowSums(.SD, na.rm=T), .SDcols = c("GB", "IE", "SEA")]
  }
  kbl(dt_sector, caption = paste0(params$species, " emissions by sector for top ten ISO totals (Gg). ISO columns are ordered largest to smallest (ISO coding = ",params$global_iso,")."))
} else {
  kbl(data.table(Note = "No sectoral NetCDF output summary found."))
}
```

```{r global-timeseries-title, echo = FALSE, results = "asis"}
if (params$domain == "GLOBAL") {
  cat("## Global Inventory Time Series\n\n")
} else {
  cat("## Inventory Time Series\n\n")
}
```

```{r ts-plots, echo = FALSE, fig.align = "center", out.width = "72%", fig.cap = "Global inventory total over time."}
if (params$domain == "GLOBAL") {
  include_plot(plot_paths$timeseries_total)
}
```

```{r ts-sector-plot, echo = FALSE, fig.align = "center", out.width = "75%", fig.cap = "Global inventory sector totals over time."}
if (params$domain == "GLOBAL") {
  include_plot(plot_paths$timeseries_sector)
}
```

```{r ts-table, echo = FALSE, results = "asis"}
if (params$domain == "GLOBAL") {
  ts <- qaqc_v2_global_timeseries(params$data_source, params$inv, params$species, params$y, years_back = 10)
  if (nrow(ts$data) > 0) {
    dt_tot <- copy(ts$data)
    dt_tot[, emis_Tg := round(emis_Tg, 3)]
    dt_sec <- copy(ts$sector)
    dt_sec[, emis_Tg := round(emis_Tg, 3)]
    dt_sec <- dcast(dt_sec, Year ~ GNFR, value.var = "emis_Tg", fill = 0)
    if ("V1" %in% names(dt_sec)) {
      data.table::setnames(dt_sec, "V1", "Other (discard)")
    }
    dt <- merge(dt_tot, dt_sec, by = "Year", all.x = TRUE)

    kbl(dt, caption = paste0("Global ", params$species, " inventory totals from ", basename(ts$file), " (Tg)"))
  } else {
    kbl(data.table(Note = "No global time-series data found."))
  }
}
```

## Monthly Emissions

```{r monthly-table, echo = FALSE, results = "asis"}
dt_out <- qaqc_v2_read_csv(summary_paths$ncoutput)
time_cols <- grep("^t[0-9]+$", names(dt_out), value = TRUE)
if (nrow(dt_out) > 0 && length(time_cols) > 1) {
  area_col <- intersect(c("Area", "ISO", "iso_char"), names(dt_out))[1]
  if (!is.na(area_col)) data.table::setnames(dt_out, area_col, "Area")
  id_cols <- setdiff(names(dt_out), time_cols)
  dt <- melt(dt_out, id.vars = id_cols, measure.vars = time_cols, variable.name = "time", value.name = "emis_t")
  dt <- dt[, .(emis_kt = round(sum(emis_t, na.rm = TRUE) / 1000, 3)), by = .(Area, time)]
  dt_wide <- dcast(dt, Area ~ time, value.var = "emis_kt", fill = 0)
  kbl(dt_wide, caption = paste0("Monthly ", params$species, " NetCDF output emissions (kt)"))
} else {
  kbl(data.table(Note = "No monthly columns available. Annual-only NetCDF summaries contain t1 only."))
}
```

## Sector Classifications

```{r sector-lookup, echo = FALSE, results = "asis"}
if (!is.null(params$dt_sec) && nrow(params$dt_sec) > 0) {
  dt <- as.data.table(params$dt_sec)
  dt[, names(dt) := lapply(.SD, function(x) if (is.character(x)) iconv(x, from = "", to = "UTF-8", sub = "") else x)]
  kbl(dt, caption = "Sectoral lookup table")
} else {
  kbl(data.table(Note = "No sector lookup table supplied."))
}
```

```{r iso-processing-title, echo = FALSE, results = "asis"}
if (params$domain == "GLOBAL") {
  cat("## Appendix 1: Global ISO Processing QAQC\n\n")
} else {

}
```

```{r global-iso-processing, echo = FALSE}
if (params$domain == "GLOBAL") {
  
  dt_stage <- qaqc_v2_stage_totals(summary_paths, params$domain)
  if (nrow(dt_stage) > 0) {
    dt <- dt_stage[, .(Total_Gg = round(sum(emis_t, na.rm = TRUE) / 1000, 2)), by = .(Area, stage)]
    dtw <- dcast(dt, Area ~ stage, value.var = "Total_Gg", fill = 0)
    setcolorder(dtw, c("Area", "inventory", "processed", "netcdf_input", "netcdf_output"))

    dtw[, data_change := round(netcdf_input - inventory, 2)]
    dtw[, change_frac := round(data_change/inventory, 2)]
    dtw <- dtw[order(data_change)]
    
    dtbl(
      dtw,
      caption = "ISO-level conservation check",
      scroll_y = "550px"
    )
  } else {
    kbl(data.table(Note = "No ISO processing-stage summaries found."))
  }

} else {
  
}
```

```{r product-process-data-change, echo = FALSE, results = "asis"}
if (params$domain == "GLOBAL") {
  # this is where we can add some code to show what data is lost when processing the global inventory, e.g. AWB
  cat("## Appendix 2: Inventory to model data change\n\n")
} else {
  
}
```